### 2023-04-13 latex/CommStyle.tex
included testimonia

\makepagestyle{HPed}
\makeoddhead{HPed}{\small\texteng{HP Transl. \& Comm.}}{}{\small\texteng{\today}}
\makeevenhead{HPed}{\small\texteng{HP Transl. \& Comm.}}{}{\small\texteng{\today}}
\makeoddfoot{HPed}{}{\small\texteng{\thepage}}{}
\makeevenfoot{HPed}{}{\small\texteng{\thepage}}{}

\SetTEIxmlExport{autopar=false}
\NewDocumentEnvironment{translation}{O{}}{\textcolor{blue}{\textbf{Transl.:}}}{}
\NewDocumentEnvironment{philcomm}{O{}}{
	\textcolor{blue}{\textbf{Comm.:}}}{}
\NewDocumentEnvironment{sources}{O{}}{
	\textcolor{blue}{\textbf{Sources:}}\linebreak}{}
\NewDocumentEnvironment{testimonia}{O{}}{
	\textcolor{blue}{\textbf{Testimonia:}}\linebreak}{}
%\newcommand{\var}[1]{\footnotesize\textup{#1}}
\def\vl{\textit{v.l.}}
\def\var#1{\\ {\small #1}}
\def\sl#1{\emph{#1}}

\linespread{1}
\setlength{\parskip}{0.3em}
\setlength\parindent{0pt}
\setlength{\vindent}{0pt}
\def\startverse{\begin{ekdverse}\begin{itshape}} %%% marked for change 2023-04-13 mm
\def\endverse{\normalsize\end{itshape}\end{ekdverse}} % \textup %%% marked for change 2023-04-13 mm
\NewDocumentEnvironment{versinnote}{O{}}{\begin{ekdverse}\begin{itshape}}{\end{itshape}\end{ekdverse}} %%% change 2023-04-13 mm
\setvnum{}