### 2023-12-02 latex/HP3_comm.tex
adapted to new input files and cleanup

\input{Preamble-Print}
\input{CommStyle}
\begin{document}
\pagestyle{HPed}
\begin{ekdosis}
\SetLineation{lineation = none,}

\chapter*{Translation \& philological commentary}
%%%%%%%%%%
\subsection*{3.1}
\begin{translation}[hp03_001]
test.
\end{translation}

\end{ekdosis}
\end{document}