### 2024-03-19 latex/HP4_comm.tex

\input{PreambleComm}
\begin{document}
\pagestyle{HPed}
\begin{ekdosis}
\SetLineation{lineation = none,}

%\chapter*{Translation \& philological commentary}

%%%%%%%%%%
\subsection*{4.0*1}
\begin{translation}[hp04_000_1] 
Homage to the guru, Śiva, who consists of \emph{nāda}, \emph{bindu} and \emph{kalā}. [The yogi] who is constantly devoted to him attains the untainted state (\emph{nirañjanapada}).
\end{translation}

%\begin{sources}[hp04_000_1]
%\end{sources}

\begin{testimonia}[hp04_000_1]
\emph{Haṃsavilāsa} 14 (p. 47)
\begin{versinnote}
\tl{namaḥ śivāya gurave nādabindukalātmane |\\+}
\tl{nirañjanapadaṃ yāti yatra yogī parāyaṇaḥ iti dhyānam || 14 ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_000_1]
%The compound \emph{yatnaparāyaṇa} is rather unusual and generally not attested elsewhere (the only exception known  to us is the \emph{Amarasiddhi}). The reading  \emph{yogaparāyaṇa} in some secondary manuscripts makes much better sense (i.e., 'one devoted to yoga') but is likely a substitution for a more difficult reading.
%adopt yatraparāyaṇaḥ [MD: done], and assume tasmai in 4.1ab

%Look into referent of nirañjana in these texts. Judit: best to take it generally (like nirāmaya, etc.).
Verses 4.0*1–4.0*16 are omitted by the \textalpha\ group and are likely not to be original. The first additional verse resembles a \emph{maṅgala} verse that one might expect to see at the beginning of a text. The second is a verse from the \emph{Gorakṣaśataka} that introduces the topic of \emph{samadhi}. Some manuscripts of the \textbeta, \textepsilon\ and \texteta\ groups have the two verses on the synonyms of \emph{rājayoga} here (on their position in the \textalpha\ group and other manuscripts, see the note to 4.29). The rest of the additional verses (4.0*5–4.0*16) are a motley collection on \emph{samādhi}, \emph{rājayoga}, the importance of the guru, dissolving the breath, \emph{suṣumṇā}, etc. In contrast to this, the \textalpha\ group begins with a cohesive discusson on absorption (4.1–3) that transitions to the gaze (4.4) and a brief discussion of \emph{śāmbhavī} and \emph{khecarī mudrā}s (4.5–4.8). The main topic of the chapter, which is meditating on the internal sound (\emph{nādānusandhāna}), begins at 4.9 in the \textalpha\ group (whereas in other groups it begins after fifty or so verses). The emphasis on \emph{nādānusandhāna} in the fourth chapter of the \textalpha\ group is consistent with the statement in verse 1.56 that \emph{nādānusandhāna} is the fourth component of Haṭhayoga.

Its likely that the term \emph{nirañjanapada} was understood here as \emph{samādhi} because \emph{nirañjana} is included in a list of synonyms of \emph{samādhi} later in this chapter (4.29).

The triad \emph{nāda}, \emph{bindu} and \emph{kalā} occurs in earlier works, in particular Śaiva Tantras, where it appears in contexts of enunciating mantras (\emph{mantroccāra}, e.g., \emph{Kubjikāmatatantra} 7.65, \emph{Jñānārṇavatantra} 2.4, \emph{Īśānaśivagurudevapaddhati} 18-110, etc.) and sometimes qualifies dieties (e.g., \emph{Parākhyatantra} 5.156ab) and gurus (e.g., \emph{Gurugītā} 64). The context can change the meaning of these terms, so we have chosen not to translate them. For a discussion of their various meanings, see \emph{Tāntrikābhidhānakośa} 2004 vol. 2: 68–73, 2013 vol. 3, 277–279.

%nāda-bindu-kalā, referring to inner sound, generative fluid, and digits of the moon (or unmanīkalā) in the context of HY? Judit: kalā is like śakti (initial source)?
%Refer to TĀK

\end{philcomm}

\begin{metre}[hp04_000_1]
Anuṣṭubh (a: na-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.0*2}
\begin{translation}[hp04_000_2]
So now I will teach the best way to \emph{samādhi}.It destroys death, has an easy method and brings about the bliss of Brahman.
\end{translation}

\begin{sources}[hp04_000_2]
\emph{Gorakṣaśataka} 64
\begin{versinnote}
\tl{athedānīṃ pravakṣyāmi samādhikramam uttamam | \\+}
\tl{mṛtyughnaṃ tu sukhopāyair brahmānandakaraṃ sadā ||\\+}
\tl{\var{64c tu sukhopāyair ] T; sukhadopāyaṃ GU}\\!}
\end{versinnote}
\end{sources}

%\begin{testimonia}[hp04_000_2]
%\end{testimonia}

%\begin{philcomm}[hp04_000_2]
%It appears that the \emph{Gorakṣaśataka}'s verse was changed by whoever added it to reflect the discursive nature of the discussion on \emph{samādhi} that follows. By changing \emph{samādhikramam uttamam} to \emph{samādhikramalakṣaṇam} (and perhaps \emph{sadā} to \emph{param}) the verse foreshadows a broader discussion of the topic samādhi. But it doesn't really work, so adopt uttamam. [JB: We have uttamam now, so this note is not needed]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.0*5}
\begin{translation}[hp04_000_5]
The unity of the self and mind arises in the same way that salt becomes identical with water through contact [with it]. That is called \emph{samādhi}.
\end{translation}

\begin{sources}[hp04_000_5]
\emph{Vivekamārtaṇḍa} 161
\begin{versinnote}
\tl{ambusaindhavayoḥ sāmyaṃ yathā bhavati yogataḥ |\\+}
\tl{tathātmamanasor aikyaṃ samādhiḥ so’bhidhīyate ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_000_5]
\emph{Haṭharatnāvalī} 4.1
\begin{versinnote}
\tl{salile saindhavaṃ yadvat sāmyaṃ bhavati yogavit |\\+}
\tl{tathātmamanasor aikyaṃ samādhiḥ so 'bhidhīyate ||\\!}
\end{versinnote}

\emph{Yuktabhavadeva} 11.29 (attr.~to Gorakṣanātha)
\begin{versinnote}
\tl{ambusaindhavayor aikyaṃ yathā bhavati yogataḥ |\\+}
\tl{tathātmamanasor aikyaṃ samādhir abhidhīyate ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 51.72
\begin{versinnote}
\tl{tad uktaṃ yogacandrikāyām –\\+}
\tl{ambusaindhavayor aikyaṃ yathā bhavati yogataḥ |\\+}
\tl{tathātmanasor aikyaṃ samādhiḥ sa vidhīyate ||\\+}
\tl{aikyaṃ abhinnatvam | yogataḥ yogābhyāsāt ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_000_5]
%The change \emph{salile saindhavaṃ} appears to have occured because of the verb \emph{bhajati}, which is different to the \emph{Vivekamartaṇḍa}'s reading of \emph{ambusaindhavayor aikyaṃ [...] bhavati [...]}. However, since groups 2 and 3 support \emph{salile}, it is possible that Svātmārāma accepted it.
%SS sees this verse as a nirukti of samādhi using sāmyam
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.0*6}
\begin{translation}[hp04_000_6]
The identity of the individual and universal selves is called \emph{samādhi}, in which all thoughts disappear.
\end{translation}

\begin{sources}[hp04_000_6]
\emph{Vivekamārtaṇḍa} 163
\begin{versinnote}
\tl{yat samatvaṃ dvayor atra jīvātmaparamātmanoḥ |\\+}
\tl{samastanaṣṭasaṃkalpaḥ samādhiḥ so’bhidhīyate ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_000_6]
\emph{Haṭharatnāvalī} 4.2
\begin{versinnote}
\tl{tat samatvaṃ bhaved atra jīvātmaparamātmanoḥ |\\+}
\tl{samastanaṣṭasaṃkalpaḥ samādhiḥ so 'bhidhīyate ||\\!}
\end{versinnote}

\emph{Yuktabhavadeva} 11.28 (attr.~to Gorakṣanātha)
\begin{versinnote}
\tl{yat sarvadvandvayor aikyaṃ jīvātmaparamātmanoḥ |\\+}
\tl{samastanaṣṭasaṃkalpaḥ samādhiḥ so 'bhidhīyate ||\\!}
\end{versinnote}

\end{testimonia}

%\begin{philcomm}[hp04_000_6]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.0*7}
\begin{translation}[hp04_000_7]
Who indeed truly knows the majesty of Rājayoga? From knowledge, liberation †arises† [and] power (\emph{siddhi}) is obtained by means of the guru's teaching.
\end{translation}

\begin{sources}[hp04_000_7]
\emph{Amanaska} 2.5
\begin{versinnote}
\tl{rājayogasya māhātmyaṃ ko vā jānāti tattvataḥ |\\+}
\tl{jñānāt siddhir muktir iti guror jñānaṃ ca labhyate ||\\+}
\tl{\var{5c jñānāt siddhir muktir iti ] Bl Jb K Ma Va:
jñānāt siddhimuktir iti Pa Pc Tr Ua Ea:
jñānasiddhir muktir iti Vb Vd:
jñānān mukteḥ siddhir iti Ja:
jñānāt sidhyati muktir hi AllN (except Na Nm Ve Ea):
jñānāt sidhyate muktir hi Ve:
jñānāt sidhyanti muktiṃ hi Na:
jñānāt sidhyati muktiṃ hi Nm:
nānāsiddhir muktir iti Tha:
jñānasiddhir bhavaty eva Mb:
jñānasiddhimuktisiddhi Pb:
jñānāt siddhimuktor iti Je}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_000_7]
\emph{Yogacintāmaṇi} f.~37v (citing the \emph{Rājayoga} [aka \emph{Amanaska}])
\begin{versinnote}
\tl{rājayogasya māhātmyaṃ ko hi jānāti tattvataḥ |\\+}
\tl{tajjñānī vasate yatra sadeśaḥ puṇyabhājanam ||\\!}
\end{versinnote}

%\emph{Yogamārgaprakāśikā} 4.11
%\begin{versinnote}
%\tl{rājayogasya māhātmyaṃ ko vā śaknoti varṇitum | \\+}
%\tl{yogasyāsya ca kartāro vijñeyās te maheśvarāḥ ||\\!}
%\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_000_7]
The third verse quarter of this verse has been subject to much revision in the \emph{Haṭhapradīpikā} and the source text, the \emph{Amanaska}. Unlike the manuscripts of the \emph{Amanaska}, those of the \emph{Haṭhapradīpikā} transmit \emph{sthitiḥ} or \emph{sthite} after \emph{muktiḥ} or \emph{mukti}. We are unable to make sense of \emph{sthiti} here, and so have cruxed this verse quarter and emended to \emph{muktiḥ sthitā} (`liberation has happened').
%?? JB. Rather than emend to muktiḥ sthitā, I think we should consider muktisthitiḥ (etaTwo and deltaOne), which is similar to jīvanmuktasthitiṃ in the Mokṣopāya, LYV, MSS, etc. We could translate it as Sternbach below ('state of liberation'). JM: I also don't think we should emend. muktisthitiḥ is possible (but MU parallel not great because of mukti vs mukta). We could also consider muktiḥ sthitiḥ “liberation [and] stability”
% Mokṣopāya
% jīvanmuktasthitiṃ brahman kṛtvā rāghavam āditaḥ|\\
% kramāt kathaya me nityaṃ bhaviṣyāmi sukhī yathā||\,1||
%  "jīvanmuktasthitiṃ" jīvanmuktamaryādāṃ
% Yogavāsiṣṭha
% vicārajā nijā jīvanmuktasthitir ihoditā |
% muninābhyāsahīnasya vyādhasya tvanavasthitiḥ || VII. 
% MSS_9744-1 
% kāryākārye kim api satataṃ naiva kartṛtvam asti  
% jīvanmuktasthitir avagato dagdhavastrāvabhāsaḥ /<BR>
% There is not always an agent in every action, good or bad ; for, the state of a jivanmukta [one who is released from bondage while living in the world] is considered to be similar to that of the splendour of a burnt cloth [which can neither be clean nor dirty]. (Sternbach's translation)
\end{philcomm}

%%%%%%%%%%
\subsection*{4.0*8}
\begin{translation}[hp04_000_8]
Letting go of sense objects, seeing the truth, [and] realising the innate state are difficult without the compassion of a good guru.
\end{translation}

%\begin{sources}[hp04_000_8]
%\end{sources}

%\begin{testimonia}[hp04_000_8]
%\end{testimonia}

%\begin{philcomm}[hp04_000_8]
%No source?
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.0*9}
\begin{translation}[hp04_000_9]
When the primal \emph{śakti} (i.e.~\emph{kuṇḍalinī}) has been woken up by means of the various postures, retentions and wonderful techniques [i.e.~\emph{mudrā}s], the breath dissolves into the void.
\end{translation}

%\begin{sources}[hp04_000_9]
%\end{sources}

\begin{testimonia}[hp04_000_9]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{haṭhapradīpikāyām—\\+}
\tl{vividhair āsanaiḥ kumbhair vicitrakaraṇair api |\\+}
\tl{prabuddhāyām ādiśaktau prāṇaḥ śūnye vilīyate ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_000_9]
In \emph{Jyotsnā} 4.10, Brahmānanda understands `the void' (\emph{śūnya}) as the central channel. In \emph{Haṭhapradīpikā} 3.4, \emph{śūnyapādavī} is a synonym of \emph{suṣumnā}.
\end{philcomm}
%?? MD: vilīyate or pralīyate?

\begin{metre}[hp04_000_9]
Anuṣṭubh (c: ra-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.0*10}
\begin{translation}[hp04_000_10]
For the yogi whose \emph{kuṇḍalinī} has awakened and who has given up all activity, the innate state automatically arises.
\end{translation}

%\begin{sources}[hp04_000_10]
%\end{sources}

\begin{testimonia}[hp04_000_10]
\emph{Yogacintāmaṇi} f.~9r (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{utpannaśaktibodhasya tyaktaniḥśeṣakarmaṇaḥ |\\+}
\tl{yoginaḥ sahajāvasthā svayam eva prajāyate ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_000_10]
%\end{philcomm}

%?? MD: prajāyate or prakāśate?

%%%%%%%%%%
\subsection*{4.0*11}
\begin{translation}[hp04_000_11]
When the breath is flowing in the central channel and the mind enters the void, the knower of yoga uproots all [past] actions.
\end{translation}
%?? JB: all [fruits of past] actions

\begin{sources}[hp04_000_11]
\end{sources}

\begin{testimonia}[hp04_000_11]
\emph{Yogacintāmaṇi} f.~9r (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{suṣumṇāvāhini prāṇe śūnye viśati mārute |\\+}
\tl{tathā samastakarmāṇi nirmūlayati yogavit ||\\!}
\end{versinnote}

\emph{Upāsanāsārasaṅgraha} p.~66 (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{suṣumṇāvāhini prāṇe śūnye viśati mānase |\\+}
\tl{tadā samastakarmāṇi nirmūlayati marmavit ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_000_11]
%\emph{tathā} may connect the two locative absolutes.
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.0*12}
\begin{translation}[hp04_000_12]
Homage to you, the immortal one. You have slain even death, into whose mouth this world, with everything that is moving and unmoving, has fallen.
\end{translation}

\begin{sources}[hp04_000_12]
\end{sources}

\begin{testimonia}[hp04_000_12]
\end{testimonia}

%\begin{philcomm}[hp04_000_12]
%It appears that this verse, which is in manuscripts of the \textbeta\ and \textepsilon\ groups, has become corrupt in manuscripts of \textgamma\, perhaps because of eye-skip (i.e., \emph{amano nirmanāḥ śūnyaṃ jagad etac carācaram}).
%?? JB should we include this note on gamma's omission? Is it necessary? JM: no, I've percented it

%JT: alt 3 is original, makes best sense, others lost pādas b and c and tried to correct. Adopt. [MD: done]
%Eye skip might explain how amarāya namas became amano nirmanaḥ śūnyaṃ jagad ... and the tadāmaraugha was dropped because it was difficult to understand or was corrupt.
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.0*13}
\begin{translation}[hp04_000_13]
When the mind has obtained equanimity and the breath is moving into the central channel, then the \emph{vajrolī} of the lineage of immortals arises even for the person whose hope is for that.
\end{translation}
% ?? JB: we didnt translate this verse in the workshop. I've adapted my translation from a similar verse in the Amaraugha. JM: equanimity is tautologous with mind, “When equanimity has been obtained"? Also I've changed tr. to reflect tadā.

\begin{sources}[hp04_000_13]
\emph{Amaraugha} 7
\begin{versinnote}
\tl{citte tu sattvam āpanne vāyau vrajati madhyame |\\+}
\tl{tadāmaraughavajrolis tadāśājīvitasya ca ||\\+}
\tl{\var{citte tu sattvam ] $\beta$ : citte samatvam Ba Ad Ed : cittau samatvam Gb Ta.
tadāmaraughavajrolis ] \textit{conj.} :
tadāmaraughavajroli Ga :
tadāmaraudhavajrolī Ae :
eṣāmarauli vajroli Gb :
eṣāmarāli vajroli Ba :
eṣāmarauli vajrauli Ta :
yeṣāmarāli vajrauli Ad :
eṣāmarolī vajrolī Ed.
tadāśājīvitasya ca ] $\beta$ :
tadā meni mameti ca Ba :
tadā mati mateti ca Ed :
mameti ca mameti ca Ad :
tadā meti mameti mateti ca Gb (\textit{unmetr.}) :
mateti mame++ Ta.
}\\!}
\end{versinnote}

\end{sources}

%\begin{testimonia}[hp04_000_13]
%\end{testimonia}

\begin{philcomm}[hp04_000_13]
The second line of this verse is problematic in both the source text (the \emph{Amaraugha}) and the manuscripts of the \emph{Haṭhapradīpikā}. It seems that the \textepsilon\ group has preserved the most plausible version of this line, which is also quite close to the source text. The line was rewritten in \etaTwo\ and other manuscripts, as well as the \emph{Jyotsnā} (4.14cd), to include the two variations of \emph{vajrolī}, namely \emph{amarolī} and \emph{sahajolī}.%??JM: comment on pāda d? It's odd, worth cruxing even.
%C2 (Grp 7) tadāmarolī ... sahajolī prajāyate (=Jyotsnā)
\end{philcomm}

%%%%%%%%%%
\subsection*{4.0*14}
\begin{translation}[hp04_000_14]
How can knowledge exist, when the mind is alive, O goddess? So long as the breath lives, the mind does not die. Breath and mind: he who leads this pair to dissolution attains liberation; in no way [does any] other man.
\end{translation}

\begin{sources}[hp04_000_14]
\emph{Candrāvalokana} 6
\begin{versinnote}
\tl{jñānaṃ kuto manasi jīvati devi yāvat \\+}
\tl{prāṇo pi jīvati mano mriyate na yāvat |\\+}
\tl{prāṇo mano dvayam idaṃ vilayaṃ prayāti\\+}
\tl{mokṣaṃ sa gacchati naro na kathaṃcid anyaḥ ||\\+}
\tl{\var{devi ] 7970 4340 T00788, kūdṛṣṭi D4345 (unmetr.)}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_000_14]
\emph{Yogacintāmaṇi} f.~17r
\begin{versinnote}
\tl{atra śivavākyam |\\+}
\tl{jñānaṃ kuto manasi jīvati devi tāvat\\+}
\tl{prāṇo 'pi jīvati mano mriyate na yāvat |\\+}
\tl{prāṇo mano dvayam idaṃ vilayaṃ nayed yo \\+}
\tl{mokṣaṃ sa gacchati naro na kathañcid anyaḥ ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} f.~2r
\begin{versinnote}
\tl{jñānaṃ kuto manasi jīvati durvikalpe \\+}
\tl{prāṇe 'pi jīvati mano mṛyate na yāvat |\\+}
\tl{prāṇo mano dvayam idaṃ na vilīyate 'tra \\+}
\tl{mokṣa na gacchati naro 'tra kathaṃ cid eva ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_000_14]
%\end{philcomm}
%?? MD: adopt prayāti (beta) in Pāda C?

\begin{metre}[hp04_000_14]
Vasantatilakā 
\end{metre}

%%%%%%%%%%
\subsection*{4.0*15}
\begin{translation}[hp04_000_15]
Both mercury and the mind are unsteady by nature. [When either] mercury is fixed [or] the mind is fixed, nothing in the world is impossible.
\end{translation}

%\begin{sources}[hp04_000_15]
%\end{sources}

\begin{testimonia}[hp04_000_15]
\emph{Yogacintāmaṇi} f.~103v
\begin{versinnote}
\tl{śāstrāntare śivavākyam—\\+}
\tl{rasasya manasaś caiva cañcalatvaṃ samāsataḥ |\\+}
\tl{raso baddho mano baddhaṃ kiṃ na sidhyati bhūtale ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_000_15]
%LO: rasa implies bindu (and the next verse is about bindu when in the AS). But the next verse is more likely from the Rasārṇava (look at the context there).
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.0*16}
\begin{translation}[hp04_000_16]
Stabilised, mercury and breath cure disease; stilled, they automatically bring back life; and bound they bestow the ability to fly, O Bhairavī.
\end{translation}

\begin{sources}[hp04_000_16]
\emph{Rasārṇava} 1.19
\begin{versinnote}
\tl{mūrchito harati vyādhiṃ mṛto jīvayati svayam |\\+}
\tl{baddhaḥ khecaratāṃ kuryāt raso vāyuś ca bhairavi ||\\!}
\end{versinnote}

Cf.~\emph{Amṛtasiddhi} 7.7
\begin{versinnote}
\tl{mūrcchito harate vyādhiṃ baddhaḥ khecaratāṃ nayet |\\+}
\tl{sarvasiddhikaro līno niścalo muktidāyakaḥ ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_000_16]
\emph{Yogacintāmaṇi} f.~103v (attr.~to \emph{śāstrāntare śivavākyam})
\begin{versinnote}
\tl{mūrchito harate vyādhiṃ mṛto jīvayati svayam |\\+}
\tl{baddhaḥ khecaratāṃ dhatte mano vāyuś ca bhairavi ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_000_16]
The \emph{Rasārṇava} is the likely source of this verse given the vocative, \emph{bhairavi}. Similar verses are found in other yoga texts (see Mallinson and Szántó 2021: 57).

Disregarding the technical meaning of the terms, the verse plays with apparent contradictions (\emph{virodhābhāsa}) in saying literally that, when mercury and the breath are unconscious, they cure disease; when dead, they restore life and, when bound, they enable one to fly up.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.1}
\begin{translation}[hp04_001]
%Now, \emph{samādhi}–
Mind is the master of the senses; breath is the master of the mind [and] dissolution [of the mind] is the master of the breath. [The yogi] should take refuge in that master, dissolution [of the mind].
\end{translation}
%?? JB: Both \alphaOne and \alphaTwo have atha samādhiḥ as a heading before this verse. Shouldnt we include this in our edition or at least note it in the apparatus? We've added atha headings elsewhere from alpha. There's also the issue of where to place atha samādhi in relation to the greyscale verses.

%\begin{sources}[hp04_001]
%\end{sources}

\begin{testimonia}[hp04_001]
\emph{Haṭharatnāvalī} 4.4
\begin{versinnote}
\tl{indriyāṇāṃ mano nātho manonāthas tu mārutaḥ |\\+}
\tl{mārutasya layo nāthaḥ sa layo nādam āśritaḥ ||\\!}
\end{versinnote}

\emph{Yogacintāmaṇi} f.~23r
\begin{versinnote}
\tl{haṭhapradīpikāyām—\\+}
\tl{indriyāṇāṃ mano nātho manonāthaś ca mārutaḥ |\\+}
\tl{mārutasya layo nāthaḥ sa layo nādam āśritaḥ ||\\!}
\end{versinnote}

\end{testimonia}

%\begin{philcomm}[hp04_001]
%
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.1*1}
\begin{translation}[hp04_001_1]
Whether or not this [dissolution] is called liberation  in another teaching, there is no difference in the bliss arising from the dissolution of mind and breath.
\end{translation}
%?? JB: 'in another doctrine' for matāntare?
%?? JB: extraordinary bliss (kaścid?)
%\begin{sources}[hp04_001_1]
%\end{sources}

\begin{testimonia}[hp04_001_1]
\emph{Yogacintāmaṇi} f.~23r (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{ayam eva tu mokṣākhyo astu vāpi matāntare |\\+}
\tl{manaḥprāṇalayo nādo na ca kaś cit vibhidyate ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_001_1]
Some of the \textbeta, \textepsilon\ and \texteta\ groups have a different version of the second half of this verse, the most plausibe of which is, 'the extraordinary bliss of dissolution of mind and breath arises in me' (\emph{manaḥprāṇalayānando mayi kaścit pravartate}). The first person pronoun suggests that this verse derives from a dialogue of some kind before it was changed to the version in the \textgamma\ and \textdelta\ manuscripts.  
\end{philcomm}

% ?? MD: comment on the variant reading "manaḥprāṇalayānando mayi kaścit pravartate"?
% JB: maybe this is the lectio difficilor and the first person pronoun was later written out.

%%%%%%%%%%
\subsection*{4.2}
\begin{translation}[hp04_002]
A motionless and unchanging dissolution [of mind], in which inhalation and exhalation have disappeared [and] perception of sense objects has ceased, is supreme for yogis.%??JM: or "The yogis’ motionless and unchanging dissolution... is the best."
\end{translation}
% on the translation of jayati as an intransitive verb, see Alexis' footnote on jayati in his forthcoming book on the Tantrāloka āhnika 1

\begin{sources}[hp04_002]
\emph{Amanaska} 2.21
\begin{versinnote}
\tl{pranaṣṭocchvāsaniśvāsaḥ pradhvastaviṣayagrahaḥ |\\+}
\tl{niśceṣṭo nirgatārambho hy ānandaṃ yāti yogavit ||\\+}
\tl{\var{pranaṣṭocchvāsa°  ] Jb Pa Va Bl Ja Ad AllSI  AllN, 
pranaṣṭasvāsa°  Mb,
pranaṣṭāsvāsa° Ua,
praṇaśocchvāsa° Ma,
pranaṣṭaḥ svāsa° Je Vb,
pranaṣṭo śvāsa°ḥ Pc,
prāṇastho śvāsa° Pb,
pranaṣṭe svāsa° K}\\!} % retained because we've read against alpha.??jM: simplify? It's only the ucch that's different. JB: done.
%\tl{\var{hy ānandaṃ yāti yogavit ]  AllNI (except  Ma Pa Pb Pc Tr Va),
%hy ānandaṃ yāti tattvavit Pa Tr Va Ea,
%svānandaṃ yāti yogavit Vd,
%hy ānandayati yoginaḥ Ma Tha,
%hy ānanday[a]ti yogī kaḥ Pc,
%hy ānanda iva yoginaḥ AllN (except  Na Nm Ea Ve),
%hy ānanda iva yogi Na Nm,
%hy ānandaś ca yoginaḥ Ve,
%hy ānandasukhayogavit Pb,
%layo jayati yoginaḥ Ad,
%layo jayati yoginām Cc}\\!} %retained to show that jayati was in the South Indian transmissin of the Amanaska.??JM: but it's in almost all HP witnesses too so perhaps variants not needed here, just mention it in comm? Or simplify? JB done
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_002]
\emph{Yogacintāmaṇi} f.~27v (attr.~to the \textit{Rājayoga})
\begin{versinnote}
\tl{praṇaṣṭocchvāsaniḥśvāsa[ḥ] vidhvastaviṣayagrahaḥ |\\+}
\tl{niśceṣṭo nirgatārambho hy ānandayati yogikaḥ ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} f.~117 (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{pranaṣṭaśvāsaniḥśvāsaḥ pradhvastaviṣayajvaraḥ |\\+}
\tl{niśceṣṭo nirvikāraś ca layo jayati yoginā ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_002]
%Adopt alpha reading. [MD: Done.]
The fourth verse quarter of this verse differs from that of the North Indian recension of the \emph{Amanaska} (quoted as the source) but is found in two South Indian manuscripts of the \emph{Amanaska} (Ad and Cc).
\end{philcomm}

%%%%%%%%%%
\subsection*{4.3}
\begin{translation}[hp04_003]
May an extraordinary absorption prevail, in which all volition has been cut off and all activity ceased, intelligible [only] by means of itself and ineffable.%??JM: which is intelligible > intelligible, and other tweaks. Prevail not great but can't think of anything else. 
%% My translation in the Amanaska (reading jāyate): [This] extraordinary absorption by which all volition has been cut off and in which all movement has ceased, is intelligible [only] by means of itself and is beyond the scope of words.
% According to the Amanaska's transmission, jayatāṃ (and its similar readings) arose at a later time.  
\end{translation}

\begin{sources}[hp04_003]
\emph{Amanaska} 2.22
\begin{versinnote}
\tl{ucchinnasarvasaṅkalpo niḥśeṣāśeṣaceṣṭitaḥ |\\+}
\tl{svāvagamyo layaḥ ko 'pi jāyate vāgagocaraḥ ||\\+}
\tl{\var{jāyate ] AllNI (except  Ma Pc), AllSI (except Ad Tha) Ea, jayatāṃ Ma Ad Nb Eb, jayatā Pc, jñāyatāṃ Tha, jagatāṃ AllN (except Nb Nu Ea Eb), japatāṃ Ca, layatāṃ Nu}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_003]
\emph{Yogacintāmaṇi} f.~27v (attr.~to the \textit{Rājayoga})
\begin{versinnote}
\tl{ucchinnasarvasaṃkalpo nirgatāśeṣaceṣṭakaḥ |\\+}
\tl{svāvagamyo layaḥ ko 'pi jayatāṃ vāgagocaraḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_003]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.4}
\begin{translation}[hp04_004]
Dissolution, which is eternal in the elements and senses, takes place where the gaze is. The gaze that has dissolved into the focus becomes the power of living beings.%%??JM: changed meditative focal point to focus (as tr. in v.6)
%JB 'occurs' for 'takes place'
\end{translation}

\begin{sources}[hp04_004]
\emph{Candrāvalokana} 8cd–9ab
\begin{versinnote}
\tl{yatra dṛṣṭir manas tatra bhūtendriyasanātanaṃ |\\+}
\tl{sā śaktis sarvabhūtānāṃ dṛṣṭir lakṣye layaṃ gatā ||\\+}
\tl{\var{sā śaktis ] 4340 4345 T00788, sa śāntis  7970}\\!}
\end{versinnote}

\emph{Kaulajñānanirṇaya} 3.2
\begin{versinnote}
\tl{yatra dṛṣṭir manas tatra bhūtendriya sapudgalaḥ \\+} 
\tl{svaśaktijīvabhūtā hi dṛṣṭilakṣair layaṃ gatāḥ \\+}
\tl{°bhūtā hi ] conj. Hatley, °bhūtāni ABCed\\!} 
% ?? Consult Shaman's critical edition.
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_004]
\emph{Yogacintāmaṇi} f.~24r
\begin{versinnote}
\tl{rājayoge—\\+}
\tl{yatra dṛṣṭir layas tatra bhūtendriyasanātanī |\\+}
\tl{syāc chaktiḥ sarvabhūtānāṃ dṛṣṭir lakṣeṇa saṃgatā ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} f.~128r
\begin{versinnote}
\tl{tad uktaṃ haṭhapradīpikāyām– \\+}
\tl{yatra dṛṣṭir layas tatra bhūtendriyasanātanī |\\+}
\tl{syāc chaktiḥ sarvabhūtānāṃ dṛṣṭir lakṣeṇa saṃgatā ||\\+}
\tl{\var{syāc chaktiḥ sarva°] B220, syārā śakti serva° 2244}\\+}
\tl{\var{lakṣeṇa saṃgatā ] 2244, lakṣyakṣayaṃ gatā B220}\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 49.25
\begin{versinnote}
\tl{tathā coktaṃ haṭhapradīpikāyām –\\+}
\tl{yatra dṛṣṭir layas tatra bhūtendriyasanātanaḥ |\\+}
\tl{na sā śaktir bhāvabhūtānāṃ dṛṣṭe lakṣe kṣayaṃ gatā ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_004]
%Adopt dṛṣṭir lakṣye layaṃ gatā. [MD: Done.]
%PYŚ: 2.18 on bhūtendriya and drś
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.5}
\begin{translation}[hp04_005]
The Vedas, Shastras and Puranas are like common prostitutes. \emph{Śāmbhavī mudrā} is unique and guarded like a woman of a good family.
\end{translation}

\begin{sources}[hp04_005]
\emph{Amanaska} 2.9
\begin{versinnote}
\tl{vedaśāstrapurāṇāni sāmānyagaṇikā iva |\\+}
\tl{ekaiva śāmbhavī mudrā guptā kulavadhūr iva ||\\!}
%\tl{\var{vedaśāstrapurāṇāni ] AllNI (except Jb Mb) AllSI AllN (except Nb Nu), vedaśāstrapurāṇādyāḥ Mb, vedaśāstrāṇi sarvāṇi Jb: vedaśāstrapurāṇādi Nb Nu}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_005]
\emph{Haṭharatnāvalī} 4.27
\begin{versinnote}
\tl{vedaśāstrapurāṇāni sāmānyagaṇikā iva |\\+}
\tl{ekaiva śāmbhavī mudrā guptā kulavadhūr iva ||\\!}
\end{versinnote}

\emph{Yogacintāmaṇi} f.~24v (attr.~to the \textit{Rājayoga})
\begin{versinnote}
\tl{vedaśāstrapurāṇaughāḥ sāmānyagaṇikā iva |\\+}
\tl{ekaiva śāmbhavī mudrā guptā kulavadhūr iva ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 49.26 (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{vedaśāstrapurāṇāni sāmānyagaṇikā iva |\\+}
\tl{ekaiva śāmbhavī mudrā sarvatantreṣu gopitā || 49.26\\!}
\end{versinnote}

%\emph{Vimalāgama} 73-005
%\begin{versinnote}
%\tl{skānde-\\+}
%\tl{vedaśāstrapurāṇādyāḥ aṣṭā veśy[ā]ṅganā iva |\\+}
%\tl{yā punaḥ śāṅkarī vidyā guptā kulavadhūr iva ||\\!}
%\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_005]
See Birch 2013:286 for more parallel verses.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.6}
\begin{translation}[hp04_006]
The focus internal, the gaze external, unblinking: this is the \emph{śāmbhavī mudrā} concealed in all the Tantras.%??JM: tr. tweaked
% JB my preference would be: The focus is internal [yet] the gaze is external and free from blinking.  
\end{translation}

\begin{sources}[hp04_006]
\emph{Amanaska} 2.10
\begin{versinnote}
\tl{antar lakṣyaṃ bahir dṛṣṭir nimeṣonmeṣavarjitā |\\+}
\tl{eṣā hi śāmbhavī mudrā sarvatantreṣu gopitā ||\\+}
\tl{\var{antar lakṣyaṃ ] AllNI (except Ja K Pb Tr Ua Vb) Tha Vd, antarlakṣya Ad Cc K: antarlakṣaṃ Tr Nw Ve: antarlakṣā Ja Ua Ea Eb, antarlakṣo Na Nq Nu, antarlakṣyo Nb Nm: antaryogaṃ Vb: antarlīno Pb}\\!}%?? JM: delete variants?
% JB: the lakṣyaṃ/lakṣa spelling is an issue in both the Amanaska and HP mss (inlcuding alpha).
\end{versinnote}

\emph{Candrāvalokana} 1
\begin{versinnote}
\tl{antar lakṣyaṃ bahir dṛṣṭir nimeṣonmeṣavarjitaḥ |\\+}
\tl{iyaṃ sā śāṃbhavī mudrā sarvatantreṣu gopitā |\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_006]
\emph{Yogacintāmaṇi} f.~24v (attr.~to the \textit{Rājayoga})
\begin{versinnote}
\tl{antar lakṣyaṃ bahir dṛṣṭir nimeṣonmeṣavarjitā |\\+}
\tl{eṣā tu śāmbhavī mudrā sarvaśāstreṣu gopitā ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_006]
See Birch 2013: 287 for more parallel verses.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.7}
\begin{translation}[hp04_007]
When the yogi's mind and breath have dissolved in the internal focus while he is looking outwards and down with a gaze in which the pupils are unmoving, even though he's not looking [at anything], this indeed is \emph{khecarīmudrā}. O guru, it manifests because of your favour and is that reality which is Śambhu's state, free from what is void and not void.
\end{translation}

\begin{sources}[hp04_007]
\emph{Candrāvalokana} 2
\begin{versinnote}
\tl{antarlakṣyavilīnacittapavano yogī yadā vartate \\+}
\tl{dṛṣṭyā niścalatārayā bahir adhaḥ paśyan na paśyet sadā |\\+}
\tl{mudreyaṃ khalu khecarī bhavati sā yuktaprasādāt guroḥ\\+}
\tl{śūnyāśūnyavivarjitaṃ sphurati yat tattvaṃ padaṃ śāṃbhavaṃ ||\\+}
\tl{\var{°pavano ] 4340 4345 T00788, °pavane 7970}\\+}
\tl{\var{sā yukta°] 4340 4345, sā yuktā  T00788, satyuṣṭa° 7970}\\+}
\tl{\var{prasādād guroḥ ] 4340 4345 T00788, prasādāt guro 7970}\\+}
\tl{\var{yat tattvaṃ ] 7970 4340 4345, cet tatvaṃ T00788}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_007]
\emph{Yogacintāmaṇi} f.~24v
\begin{versinnote}
\tl{haṭhapradīpikāyām—\\+}
\tl{antarlakṣyavilīnacittapavano yogī yadā vartate\\+}
\tl{dṛṣṭyā niścalatārayā bahir asau paśyann apaśyann api |\\+}
\tl{mudreyaṃ khalu śāṃbhavī bhavati sā yuṣmatprasādād guro\\+}
\tl{śūnyāśūnyavivarjitaṃ sphurati yat tatvaṃ padaṃ śāṃbhavam ||\\!}
\end{versinnote}

\emph{Anubhavanivedana} 1
\begin{versinnote}
\tl{antarlakṣyavilīnacittapavano yogī yadā vartate\\+}
\tl{dṛṣṭyā niścalatārayā bahir asau paśyann apaśyann api |\\+}
\tl{mudreyaṃ khalu śāmbhavī bhavati sā yuṣmatprasādād guro\\+}
\tl{śūnyāśūnyavivarjitaṃ bhavati yat tattvaṃ padaṃ śāmbhavam ||1 ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 49.27 (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{antarlakṣavilīnacittapavano yogī yadā varttate\\+}
\tl{dṛṣṭyā niścalatārayā bahir adhaḥ paśyann apaśyann api |\\+}
\tl{mudreyaṃ khalu śāmbhavī bhavati sā yuṣmatprasādāt guroḥ\\+}
\tl{śūnyāśūnyavivarjitaṃ sphurati yat tatvaṃ padaṃ śāmbhavam ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_007]
On the similarity of \emph{śāmbhavī} and \emph{khecarī mudrās} in 4.5–7, see the note on 4.8. 
\end{philcomm}

\begin{metre}[hp04_007]
Śārdūlavikrīḍita 
\end{metre}

%%%%%%%%%%
\subsection*{4.8}
\begin{translation}[hp04_008]
There is no difference between the states of \emph{śāmbhavī} and \emph{khecarī}.
\end{translation}
% Adopt avasthāyām abhedatā (N10, N12, T11) (Cf.~avasthāyām abhedataḥ C6)
%
% śrīśāmbhavyāś ca khecaryā avasthāyām abhedataḥ |M1,J10
% bhavec cittalayākārā unmanī citsvarūpiṇī | J10
% The beyond-mind [state], whose form is the dissolution of mind and own nature is consciousness, is identical in the state of Śāmbhavī and Khecarī.

%\begin{sources}[hp04_008]
%\end{sources}

\begin{testimonia}[hp04_008]
\emph{Yogacintāmaṇi} f.~25r (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{śrīśāṃbhavyā khecaryāś ca avasthāṃ ca labhed ataḥ |\\+}
\tl{tāre jyotiṣi saṃyojya kiñ cid uccālayed bhruvau ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} f.~128r–128v (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{tāre jyotiṣi saṃyojya kiṃ cid unnamayed bhuvau [|]\\+}
\tl{śrīśāṃbhavyāś ca khecaryā avasthā[ṃ] tu labhed ataḥ ||\\+}
\tl{\var{unnamayed ] 2244, saṃcālayed B220}\\+}
\tl{\var{unmanīkaraṇaṃ kṣaṇāt ] 2244, unmanīkāralakṣaṇam B220}\\+}
\tl{\var{tu labhed ataḥ ] 2244, labhate daśa B220}\\!}
\end{versinnote}

\end{testimonia}

\begin{philcomm}[hp04_008]
This line may be authorial because it is explaining the fact that 4.6 and 4.7 teach similar techniques called \emph{śāmbhavī mudrā} and \emph{khecarīmudrā} respectively. One would expect a verse on \emph{khecarī mudrā} to be explaining the insertion of the tongue in the nasopharyngeal cavity (cf.~3.33–48), but the practice of meditating by fixing the gaze is called \emph{khecarī mudrā} in 4.7 (as attested by \textalpha, \textbeta\ and \texteta\ manuscripts) and manuscripts of the \emph{Candrāvalokana}, the text from which Svātmārāma borrowed this verse. It is therefore likely that he added 4.8 to explain that \emph{śāmbhavī} and \emph{khecarī mudrās} are the same with respect to the gaze.

%Judit would greyscale 4.8, thinking that it is a marginal note. And greyscale the pātāle verses.
%avasthā: Brahmānanda explains this as referring to the place (dhāman) of khecarī (the middle of the brow) and śāmbhavī (the heart).
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*1}
\begin{translation}[hp04_008_1]
[The yogi] should fix the pupils on a light and slightly raise the eyebrows. This is the way of the preliminary yoga, which immediately brings about the beyond-mind state.
\end{translation}
% unmanīkārakaḥ ?

\begin{sources}[hp04_008_1]
\emph{Amanaska} 1.8 (South Indian Recension)
\begin{versinnote}
\tl{netre jyotiṣi saṃyojya kiṃ cid unnamayed bhruvau |\\+}
\tl{pūrvayogasya mārgo 'yam unmanīkārakaḥ kṣaṇāt ||\\+}
\tl{\var{unnamayed ] Cb Vd: unnamayan Bb Tha: unmīlaye W: unmilayet Uc: unmanaya Pe: unmīlya Cc}\\!}
%\tl{\var{unmanīkārakaḥ kṣaṇāt ] Bb Tha W Vd Pe: unmunikārakakṣaṇāt Uc: unmanīkāraṇaṃ kṣaṇāt Cc: unmanākāśakaḥ kṣaṇāt Ad Ca: unmanākāśakakarṣaṇāt Cb}\\!}%?? variants needed? JB I've deleted the variants for the last pāda but we need those on unnamayed (see below) 
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_008_1]
\emph{Yogacintāmaṇi} f.~25r (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{tāre jyotiṣi saṃyojya kiṃ cid uccālayed bhruvau |\\+}
\tl{pūrvayogasya mārgo 'yam unmanīkaraṇaḥ kṣaṇāt ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} ff. 128r-128v (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{tāre jyotiṣi saṃyojya kiṃ cid unnamayed bhuvau |\\+}
\tl{pūrvayogasya mārgo [']yam unmanīkaraṇaṃ kṣaṇāt ||\\!}
\end{versinnote}

%Rājayogāmṛta 3.7c-d $-$ 3.8a-b
%\begin{versinnote}
%\tl{tārāj jyotiṣi saṃyojya kiṃcid unmīlayed bhruvau |\\+}
%\tl{pūrvayogasya mārgo 'yaṃ unmanitārakaṃ kṣaṇāt ||\\!}
%\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_008_1]
%MD: Better to adopt unnamayed and kārakaḥ kṣaṇāt?
% ?? JB Since unnamayed in the beta witnesses and source, and makes better sense, I think we should adopt. The dictionary states that uc-cal means 'to go or move away from; to free or loosen one's self from', which is not so good here.

This and the next verse (4.8*1–2) are from the South Indian recension of the \emph{Amanaska}, a later rewriting of that work which incorporates these verses in an additional passage on Tārakayoga, added here to elaborate on the meditation of fixing the gaze that is taught in 4.4–7. 
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*2}
\begin{translation}[hp04_008_2]
Some are confused by a multitude of [tantric] texts, some by a mass of vedic texts and some by reasoning. They do not know what causes one to cross over.%??JM: put Tāraka and explain in note? Tārakayoga is mentioned in the previous note.
% JB in the source 
\end{translation}
% Maybe grey scale

\begin{sources}[hp04_008_2]
\emph{Amanaska} 1.11 (South Indian Recension)
\begin{versinnote}
\tl{ke cid āgamajālena ke cin nigamasaṅkulaiḥ |\\+}
\tl{ke cit tarkeṇa muhyanti naiva jānanti tārakam ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_008_2]
\emph{Yogacintāmaṇi} f.~25r (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{ke cid āgamajālena ke cin niyamasaṃkulāḥ |\\+}
\tl{ke cit tarkeṇa muhyanti naiva jānanti tārakam ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 49.29 (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{ke cid āgamajālena ke cin nigamasaṃkule |\\+}
\tl{ke cit tarkeṇa muhyanti naiva jānanti tārakam ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_008_2]
On why this verse is in greyscale, see the note to 4.8*1.
% ??JB we should adopt nigama in pāda b because it is attested by several manuscript groups and the source.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*3}
\begin{translation}[hp04_008_3]
With eyes half open, mind steady, gaze placed at the tip of the nose and inner state motionless, [the yogi] should lead the moon and sun to dissolution. He attains the state of the highest reality, the supreme essence, whose form is light and which is devoid of anything external and shining intensely. What more is to be said here?
\end{translation}

\begin{sources}[hp04_008_3]
\emph{Candrāvalokana} 3
\begin{versinnote}
\tl{ardhotghāṭitalocana[ḥ] sthiramanā nāsāgradattekṣaṇaś\\+}
\tl{candrārkāv api līnatām upagatau niṣpandarūpaṃ vapuḥ |\\+}
\tl{jyotīrūpam aśeṣabāhyarahitaṃ dedīpyamānaṃ paraṃ\\+}
\tl{tatvaṃ tatpadam eti yat tu paramaṃ vācyaṃ kim atrādhikaṃ ||\\+}
\tl{\var{sthiramanā ] 4340 4345 T00788, sphuramanā 7970}\\+}
\tl{\var{nāsāgradattekṣaṇaś ] 7970 4345 T00788, nāsāgradaṭhattakṣaṇaś 4340}\\+}
\tl{\var{tatvaṃ tatpadam eti yat tu] 4340 4345 T00788, tatvaṃ-n-tatpadam eti yat kṛ 7970}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_008_3]
\emph{Yogacintāmaṇi} ff. 24v–25r (attr.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{ardhodghāṭitalocanaḥ sthiramanā nāsāgradattekṣaṇaḥ\\+}
\tl{candrārkāv api līnatām upanayen niḥspandabhāvottare |\\+}
\tl{jyotīrūpam aśeṣabāhyarahitaṃ dedīpyamānaṃ paraṃ  \\+}
\tl{tatvaṃ tat padam eti vastu paramaṃ vācyaṃ kim atrādhikam ||\\!}
\end{versinnote}

\emph{Anubhavanivedana} 2
\begin{versinnote}
\tl{ardhodghāṭitalocanaḥ sthiramanā nāsāgradattekṣaṇaś\\+}
\tl{candrārkāv api līnatām upagatau trispandabhāvāntare |\\+}
\tl{jyotīrūpam aśeṣabāhyarahitaṃ caikaṃ pumāṃsaṃ param\\+}
\tl{tattvaṃ tatpadam eti vastu paramaṃ vācyaṃ kim atrādhikam ||2 ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 49.30 (citing the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{ardhodghāṭitalocanaḥ sthiramanā nāsāgradattekṣaṇaḥ\\+}
\tl{candrārkāvapi līnatāmupanayennispandavācyaṃ tataḥ |\\+}
\tl{jyotīrūpaviśeṣabāhyarahitaṃ dedīpyamānaṃ paraṃ\\+}
\tl{tatvaṃ tatparamasti vastu paramaṃ vācyaṃ kimatrādhikam || iti ||\\!}
\end{versinnote}

\end{testimonia}

\begin{philcomm}[hp04_008_3]
This verse is not in the \textalpha\ manuscripts and was probably added because it elaborates on the gaze. In the \emph{Candrāvalokana} (i.e., the source text) and the \emph{Anubhavanivedana}, it follows 4.7, which may explain its position in the \textgamma\ and \textdelta\ groups.
\end{philcomm}
% ?? MD: I have changed the verse order according to the epsilon group. Please adjust the commentary. JB Done

\begin{metre}[hp04_008_3]
Śārdūlavikrīḍita 
\end{metre}

%%%%%%%%%%
\subsection*{4.8*4}
%\begin{translation}[hp04_008_4]
%That which enters the aperture into the underworld, which exists at the base of Meru,  the wise yogin say that is the truth, the source of [all] rivers. The essence of the body flows from the moon. Because of that, death arises for people. [This] good mudrā should block that [essence]. Otherwise, there is no bodily perfection.
%\end{translation}
% See 3.49

%\begin{sources}[hp04_008_4]
%\end{sources}

% \begin{testimonia}[hp04_008_4]
% \emph{Haṭharatnāvalī} 4.30
% \begin{versinnote}
% \tl{pātāle yad viśati suṣiraṃ merūmūlaṃ tad asti \\+}
% \tl{tattvaṃ caitad vadati sudhā tanmukhaṃ nimnagānām |\\+}
% \tl{candrāt sāraṃ sravati vapuṣas tena mṛtyur narāṇāṃ\\+}
% \tl{tad badhnīyāt sukharatimṛdur nānyathā kāryasiddhiḥ ||\\!}
% \end{versinnote}

% \emph{Yogacintāmaṇi} f.~74v (attr.~to the \emph{Haṭhapradīpikā})
% \begin{versinnote}
% \tl{pātāle yad viyati śikhare merumūle tad asti\\+}
% \tl{tattvaṃ caitat pravadati sudhīs sammukhe nimnagānām |\\+}
% \tl{candrāt sāraṃ sravati vapuṣas tena mṛtyur narāṇāṃ\\+}
% \tl{taṃ badhnīyāt svakaraṇamṛdā nānyathā kāyasiddhiḥ ||\\!}
% \end{versinnote}

% \emph{Yuktabhavadeva} 7.220 (attr.~to Gorakṣanātha)
% \begin{versinnote}
% \tl{pātāle yad viśati suṣiraṃ merumūle yad asti \\+}
% \tl{tadvac caitat pravadati sudhīs tanmukhaṃ nimnagānām |\\+}
% \tl{candrāt sāraḥ sravati vapuṣas tena mṛtyur narāṇām \\+}
% \tl{badhnīyāttat sukaraṇam atho nānyathā kāyasiddhiḥ ||\\!}
% \end{versinnote}

% \emph{Haṭhatattvakaumudī} 14.26 (attr.~to the \emph{Haṭhapradīpikā})
% \begin{versinnote}
% \tl{pātāle yadvitatasuṣiraṃ merumūle tad asmin\\+}
% \tl{tadvac caitat pravadati sudhīs tanmukhaṃ nimnagānām |\\+}
% \tl{candrāt sāraḥ sravati vapuṣas tena mṛtyur narāṇāṃ \\+}
% \tl{taṃ badhnīyāt sukaraṇamṛdā nānyathā kāyasiddhiḥ ||\\!}
% \end{versinnote}
% \end{testimonia}

\begin{philcomm}[hp04_008_4]
For the translation and testimonia, as well as an explanation of the various places and versions of this verse in the text, see 3.49.
\end{philcomm}

%\begin{metre}[hp04_008_4]
%Mandākrāntā 
%\end{metre}

%%%%%%%%%%
\subsection*{4.8*5}
\begin{philcomm}[hp04_008_5]
See 3.75, where this verse is also found.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*6}
\begin{philcomm}[hp04_008_6]
See 3.76, where this verse is also found.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*7}
\begin{translation}[hp04_008_7]
[The yogi] should not worship the \emph{liṅga} by day, nor should he worship it by night. He should worship the \emph{liṅga} constantly, by suppressing day and night.
\end{translation}

\begin{sources}[hp04_008_7]
\emph{Khecarīvidyā} 3.19
\begin{versinnote}
\tl{na divā pūjayel liṅgaṃ na rātrau ca maheśvari |\\+}
\tl{sarvadā pūjayel liṅgaṃ divārātrinirodhataḥ ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_008_7]
\emph{Haṭhasaṅketacandrikā} f.~128v (attr.~Haṭhapradīpikā)
\begin{versinnote}
\tl{divā na pūjayel liṃga[ṃ] rātrau naiva prapūjayet [|]\\+}
\tl{satataṃ pūjayel liṃga[ṃ] divārātrau ca pūjayet [||]\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_008_7]
%?? JM: nirodhataḥ reading much better. Early corruption?
% JB I think we should note nirodhataḥ as the better reading in the comments but adopt divārātrau (or divārātraṃ) na pūjayet (he should not worship at day and night (i.e., when prāṇa is moving in the sun and mmoon channels)). 
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*8}
\begin{translation}[hp04_008_8]
Now \emph{khecarī}:\\
There is a hollow that generates knowledge and has five streams. \emph{Khecarīmudrā} is situated in that pure void.
\end{translation}
% ?? compound khecarīmudrā in 4.8*8c and 4.8*9c?

%\begin{sources}[hp04_008_8]
%\end{sources}

\begin{testimonia}[hp04_008_8]
\emph{Haṭhasaṅketacandrikā} f.~128v
\begin{versinnote}
\tl{suṣiraṃ jñānajanakaṃ paṃcastr[o]taḥsamanvitaṃ |\\+}
\tl{tiṣṭhate khecarī mudrā tasmin [ś]ūnye niraṃjane  ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_008_8]
This verse may not be referring to the cavity in which the tongue is placed but perhaps to a more esoteric sense based on the meaning of \textit{pañcasrotas} as the five streams of tantric Śaiva teachings, which in this case generate knowledge. On \textit{pañcasrotas}, see \emph{Tāntrikābhidhānakośa} 2013, vol. 3: 361.
\end{philcomm}

\begin{metre}[hp04_008_8]
Anuṣṭubh (a: na-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.8*9}
\begin{translation}[hp04_008_9]
The breath in the left and right channels moves into the middle. Without doubt, \emph{khecarīmudrā} abides in that place.
\end{translation}

%\begin{sources}[hp04_008_9]
%\end{sources}

\begin{testimonia}[hp04_008_9]
\emph{Haṭhasaṅketacandrikā} f.~??
\begin{versinnote}
\tl{savyadakṣiṇanāḍīstho madhye carati mārutaḥ |\\+}
\tl{tiṣṭhati khecarī mudrā tasmin sthāne na saṃśayaḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_008_9]
%The term \emph{madhye} appears to be referring to the central channel here. 
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*10}
%\begin{translation}[hp04_008_10]
%The mind moves in the void because the tongue moves in the void, which is why that \emph{mudrā} is worshipped by the adepts with the name “she who moves in the void” (\emph{khecarī}).
%\end{translation}

% \begin{sources}[hp04_008_10]
% \emph{Vivekamārtaṇḍa} 50
% \begin{versinnote}
% \tl{cittaṃ carati khe yasmāj jihvā carati khe gatā |\\+}
% \tl{[tenai]ṣā khecarī nāma mudrā siddhair namaskṛtā ||\\!}
% \end{versinnote}

% %\emph{Skandapurāṇa} 4.41.45%??JM: is it worth putting this in? Verse is prob found all over the place and this isn't the old SP. Just YCM will do.
% %\begin{versinnote}
% %\tl{cittaṃ carati khe yasmā jihvā carati khe gatā |\\+}
% %\tl{tenaiṣā khecarī nāma mudrā siddhair niṣevitā। ||\\!}
% %\end{versinnote}
% \end{sources}

% \begin{testimonia}[hp04_008_10]
% \emph{Yogacintāmaṇi} ?? [attrib. Skandapurāṇa]
% %\begin{versinnote}
% %\tl{cittaṃ carati khe yasmāj jihvā carati khe gatā |\\+}
% %\tl{tenaiṣā khecarī nāma mudrā siddhair niṣevitā ||\\!}
% %\end{versinnote}

% \emph{Yuktabhavadeva} 212

% \end{testimonia}

\begin{philcomm}[hp04_008_10]
See 3.37, where this verse is also found.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*11}
\begin{translation}[hp04_008_11]
At the juncture of Iḍā and Piṅgalā, the void devours the breath. \emph{Khecarīmudrā} abides there. This is true, time and time again.
\end{translation}

%\begin{sources}[hp04_008_11]
%\end{sources}

\begin{testimonia}[hp04_008_11]
\emph{Upāsanāsārasaṅgraha} (IFP Transcript T1095) p.~42.
\begin{versinnote}
\tl{iḍāpiṅgalayor yoge śūnye caivānilaṃ graset |\\+}
\tl{tiṣṭhate khecarī mudrā tatra satyaṃ punaḥ punaḥ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_008_11]
Cf.~Rāghavabhaṭṭa \emph{ad Śāradātilaka} 25.43: \emph{suṣumṇāyām eteṣu parvasu iḍāpiṅgalayor yogo bhavatīti jñeyam}.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*12}
\begin{translation}[hp04_008_12]
The \emph{mudrā} situated in the cakra of the void (\emph{vyomacakre}) in the middle of the sun and moon on an unsupported surface is [the \emph{mudrā}] called \emph{khecarī}.
\end{translation}

%\begin{sources}[hp04_008_12]
%\end{sources}

\begin{testimonia}[hp04_008_12]
\emph{Upāsanāsārasaṅgraha} (IFP Transcript T1095) p.~41.
\begin{versinnote}
\tl{somasūryadvayor madhye nirālambe tale punaḥ |\\+}
\tl{saṃsthitā vyomacakre sā mudrā nāma ca khecarī ||\\!}
\end{versinnote}

\emph{Gorakṣasiddhāntasaṅgraha} p.~37
\begin{versinnote}
\tl{sūryācandramasormadhye nirālambe'nile punaḥ |\\+}
\tl{saṃsthitā vyomacakre yā sā mudrā nāma khecarī ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_008_12]
The reading \emph{tale} is suspect, as is \emph{nirālambāntare}, the conjecture of Brahmānanda.
%  bhruvormadhye dvyasraṃ vyomacakraṃ tad dhi cinmayasya jyotiṣa upalabdhisthānam  Commentary on LYV 9.82.
% Haṭhasaṅketacandrika (4192): śūnyaṃ vyomacakravyavasthitam | iḍāpiṃgalāṃtargataṃ brahmaraṃdhrākhyaṃ vyomacakraṃ [...]. 
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*13}
\begin{translation}[hp04_008_13]
Brought forth by me, lovely, and clearly beloved of Śiva, the Suṣumṇā should fill herself with the divine air through her rear mouth.
\end{translation}

%\begin{sources}[hp04_008_13]
%\end{sources}

\begin{testimonia}[hp04_008_13]
\emph{Yogasārasaṅgraha} p. 61 (attrib. Praṇavacintāmaṇi)
\begin{versinnote}
\tl{sā mayā viditā yā māyā sākṣācchivavallabhā |\\+}
\tl{pūrayen mārutaṃ divyaṃ suṣumnā paścime mukhe ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_008_13]
The referent of \emph{mayā} is unspecified and no source text has been identified. It most probably refers to a goddess.%to understand here because the referent of \emph{mayā} is not known. If it refers to a yogi, then it might mean that \emph{suṣumnā} has been pierced. Otherwise, if \emph{udbheditā} means `brought forth', one would assume the speaker to be a god or goddess. 
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*14}
\begin{translation}[hp04_008_14]
And if [she] fills herself from the front, \emph{khecarī} definitely arises. [The yogi] should practise \emph{khecarīmudrā}. The state beyond mind arises.
\end{translation}

%\begin{sources}[hp04_008_14]
%\end{sources}

\begin{testimonia}[hp04_008_14]
\emph{Upāsanāsārasaṅgraha} p.~135
purastāc caiva pūryeta niścitā khecarī bhavet |
abhyaset khecarīmudrāmunmanī saṃprajāyate ||
\end{testimonia}

\begin{philcomm}[hp04_008_14]
We have understood \emph{niścitā} as an adverb. No witnesses have \emph{niścitam}, but we see no other way of taking \emph{niścitā}.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*15}
\begin{translation}[hp04_008_15]
[The yogi] should practise \emph{khecarīmudrā} until he falls into a yogic sleep. For one who has attained yogic sleep, death never arises.
\end{translation}

%\begin{sources}[hp04_008_15]
%\end{sources}

\begin{testimonia}[hp04_008_15]
\emph{Upāsanāsārasaṅgraha} p.~135\\
abhyaset khecarīṃ tāvat yāvat syād yoganidrataḥ |
saṃprāptayoganidrasya kālo nāsti kadācana ||
\end{testimonia}

\begin{philcomm}[hp04_008_15]
\end{philcomm}

%%%%%%%%%%

\subsection*{4.8*16}
\begin{translation}[hp04_008_16]
Between the eyebrows is the place of Śiva. The mind dissolves there. That level should be known as the fourth state. Death does not exist there.
\end{translation}
%?? JB: pada here means place/site (level sounds strange)
% abode of Śiva ... that place should be known

%\begin{sources}[hp04_008_16]
%\end{sources}

\begin{testimonia}[hp04_008_16]
\emph{Upāsanāsārasaṅgraha} p.~135
bhruvor madhye śivasthānaṃ manas tatra vilīyate |
jñātavyaṃ tatparaṃ turyaṃ tatra kālo na vidyate ||
\end{testimonia}

%\begin{philcomm}[hp04_008_16]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*17}
\begin{translation}[hp04_008_17]
Between the moon and the sun, [the yogi] should apply \emph{khecarīmudrā}, which is situated in the supportless, great void, the \emph{vyomacakra}.
\end{translation}

\begin{sources}[hp04_008_17]
\emph{Jñānasāra} 3.3cd
\begin{versinnote}
\tl{candrasūryadvayor madhye muḍādadyā? tu khecarīm |\\!}
\end{versinnote}

\end{sources}

%\begin{testimonia}[hp04_008_17]
%\end{testimonia}

\begin{philcomm}[hp04_008_17]
The \emph{vyomacakra} ('the cakra of space') is also mentioned in a half-verse that was added to some later recensions of the \emph{Haṭhapradīpikā} (see 3.35) and states that \emph{vyomacakra} is another name for \emph{khecarīmudrā}. In \emph{Jyotsnā} 4.45, Brahmānanda states that the \emph{vyomacakra} is associated with all the voids in the middle of the brow (\emph{bhrūmadhye sarvakhānāṃ samanvayāt}) and, in the \emph{Haṭhasaṅketacandrikā} (f. 129v), Sundaradeva says that it is called the Brahmarandhra, which is between Iḍā and Piṅgalā (\emph{iḍāpiṅgalāntargataṃ brahmarandhrākhyaṃ
vyomacakraṃ tat khecarīmudrāṃ śaktiṃ kuryād ...}). 
%?? Write a note about epsilon 3 vyomacakram vyavasthitaṃ: Cf.~the greyscale verse in 3.35.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*18}
\begin{translation}[hp04_008_18]
[The yogi] should make the mind supportless and think of nothing at all. He assuredly remains like a pot in the ether, [empty] inside and outside.
\end{translation}

\begin{sources}[hp04_008_18]
\emph{Jñānasāra} 3.4
\begin{versinnote}
\tl{nirālambaṃ manaḥ kṛtvā na kiñcid api cintayet || 3 ||\\+}
\tl{sa bāhyābhyantare yogī ghaṭavat tiṣṭhate priye |\\!}
\end{versinnote}

% \emph{Śivasaṃhitā} 5.210
% \begin{versinnote}
% \tl{nirālambo bhavej jīvo jñātvā vedāntayuktitaḥ |\\+}
% \tl{nirālambaṃ manaḥ kṛtvā na kiṃ cic cintayet sudhīḥ ||\\!}
% \end{versinnote}
\end{sources}

%\begin{testimonia}[hp04_008_18]
%\end{testimonia}

%\begin{philcomm}[hp04_008_18]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*19}
\begin{translation}[hp04_008_19]
When the external air has dissolved into the void, the breath is sure to go to its place †with the mind on the side of the sun†.
%When the external air has dissolved into the void, the breath surely goes to its own place, † and the sun together with the mind to the fire.†
\end{translation}

%\begin{sources}[hp04_008_19]
%\end{sources}

\begin{testimonia}[hp04_008_19]
\emph{Haṭhapradīpikā} (with ten-chapters) 7.52 
\begin{versinnote}
\tl{bāhyavāyur yathā līnaḥ svasya madhye na saṃśayaḥ |\\+}
\tl{svasthānaṃ gacchati prāṇaḥ sūryo 'gnau pavane tathā ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_008_19]
% sūryāgnir manasā tathā M1
It is hard to make sense of \emph{sūryāṅge} here (`on the side of the sun' ?). The terms \emph{sūryāṅge} and \emph{candrāṅge} occur in the third chapter (3.15) in the sense of the right and left sides of the body, respectively. However, this meaning does not seem relevant here. Since we are not sure of the meaning of \emph{sūryāṅge} and since the variant readings with \emph{pavana} are not clear either, we have cruxed the fourth verse quarter.

%\textit{Maitrī Upaniṣad} 6.1-2 might provide a clue to the meaning of \emph{sūryāṅga}. This Upaniṣad explains that the \textit{ātman} manifests itself in two ways: as \textit{prāṇa} internally and as the sun externally. \textit{Prāṇa} is identified as the inner aspect of the \textit{ātman}. It is associated with breath, etc. The sun is depicted as the external manifestation of the \textit{ātman}. It symbolizes the cosmic source of energy, etc. The text emphasizes that the movements of \textit{prāṇa} and the sun are synchronized and states that the golden \textit{puruṣa}, who dwells in the sun, also resides in the heart as \textit{prāṇa}. If one is willing to accept a connection between this verse and the concept mentioned in the \textit{Maitrī Upaniṣad}, then \textit{sūryāṅge} possibly has even two layers of meaning. On the microcosmic level, \textit{prāṇa} would return home to the heart. Moreover, on the macrocosmic level, the \textit{ātman} would return to its macrocosmic home - the sun.
% ?? Does sūryāṅge occur in the Maitrī Upaniṣad? If not, the connection with this Upaniṣad seems tenuous. 
\end{philcomm}

%?? MD: yadā (epsilon) or yathā (beta)?
% JB separate yadā līnaḥ in the edition (yadā seems easier to understand than yathā)

%%%%%%%%%%
\subsection*{4.8*20}
\begin{translation}[hp04_008_20]
For [the yogi] practising in this way day and night on the path of the breath, as a result of the practice the breath is consumed, [and] the mind dissolves into it.
\end{translation}

%\begin{sources}[hp04_008_20]
%\end{sources}

\begin{testimonia}[hp04_008_20]
\emph{Haṭhasaṅketacandrikā} f. 129v
\begin{versinnote}
\tl{abhyāsāl līyate vāyuḥ manas tatra vilīyate ||\\!}
\end{versinnote}
\end{testimonia}
%??JM: I can't find these HSC refs

\begin{philcomm}[hp04_008_20]
Brahmānanda identifies the path of the breath (\emph{vāyumārga}) with Suṣumṇā.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*21}
\begin{translation}[hp04_008_21]
[The yogi] should flood the body with nectar from the soles of the feet to the head. He is sure to become perfected, with a great body and great strength and valour.\\Thus ends \emph{khecarī}.
\end{translation}

%\begin{sources}[hp04_008_21]
%\end{sources}

\begin{testimonia}[hp04_008_21]
\emph{Haṭhasaṅketacandrikā} f. 129v
\begin{versinnote}
\tl{siddhaty evaṃ tadā kāyo mahābalaparākramaḥ ||\\!}
\end{versinnote}
\end{testimonia}
%??JM: can't find. ref?

%\begin{philcomm}[hp04_008_21]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*22}
\begin{translation}[hp04_008_22]
Now \emph{śāmbhavī}:\\
{[}The yogi] should [put] the mind in Śakti and Śakti in the mind, observe the mind with the mind, and meditate on it as the supreme state.
\end{translation}
%V3 is closer to the CA reading

\begin{sources}[hp04_008_22]
\emph{Candrāvalokana} 27
\begin{versinnote}
\tl{śaktimadhye manaḥ kṛtvā manaś śaktes tu madhyamam |\\+}
\tl{manasā mana ālokya taṃ dhyāyet paramaṃ padaṃ ||\\+}
\tl{\var{madhyamam ] madhyagaṃ 4340}\\+}
\tl{\var{taṃ dhyāyet ] tad dhyāyet 4340}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_008_22]
\emph{Haṭhasaṅketacandrikā} f.~129v–130r
\begin{versinnote}
\tl{śaktimadhyo manaḥ kṛtvā śaktiṃ ca svāntamadhyagāṃ |\\+}
\tl{manasā mana ālokya tad dhyāyet paramaṃ padam ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_008_22]
In the second verse quarter, \emph{mana} for \emph{mano} is for the metre.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*23}
\begin{translation}[hp04_008_23]
Put the self in space and put space in the self. [The yogi] should make everything consist of space and think of nothing at all.
\end{translation}

\begin{sources}[hp04_008_23]
\emph{Uttaragītā} 1.9%??JM: bibliog for texts cited — where?
\begin{versinnote}
\tl{khamadhye kuru cātmānam ātmamadhye ca khaṃ kuru |\\+}
\tl{ātmānaṃ khamayaṃ kṛtvā na kiṃ cid api cintayet  ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_008_23]
\emph{Haṭharatnāvalī} 4.45
\begin{versinnote}
\tl{khamadhye kuru cātmānam ātmamadhye ca khaṃ kuru |\\+}
\tl{sarvaṃ ca khamayaṃ kṛtvā na kiṃ cid api cintayet ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_008_23]
%?? Pāda c: sarvaṃ or ātmānaṃ?
%\end{philcomm}
% J up to here! 11 March

%%%%%%%%%%
\subsection*{4.8*24}
\begin{translation}[hp04_008_24]
Like an empty pot in air, [the yogī] is empty on the inside and empty on the outside. Like a full pot in the ocean, [the yogi] is full on the inside and full on the outside.%??JM: I changed the sky to air. Not sure, but I don't think we are meant to think of a pot floating in space.
\end{translation}

\begin{sources}[hp04_008_24]
\emph{Laghuyogavāsiṣṭha} 6.15.79 (\emph{Mokṣopāya} 6.155.25)
\begin{versinnote}
\tl{antaḥśūnyo bahiḥśūnyaḥ śūnyakumbha ivāmbare |\\+}
\tl{antaḥpūrṇo bahiḥpūrṇaḥ pūrṇakumbha ivārṇave ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_008_24]
\emph{Yogacintāmaṇi} (attrib.~Vasiṣṭha, = \emph{Laghuyogavāsiṣṭha} 15.79)
\begin{versinnote}
\tl{antaḥśūnyo bahiḥśūnyaḥ śūnyakumbha ivāmbare |\\+}
\tl{antaḥpūrṇo bahiḥpūrṇaḥ pūrnakumbha ivārṇave ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.46
\begin{versinnote}
\tl{antaḥpūrṇo bahiḥpūrṇaḥ pūrṇakumbha ivāmbhasi |\\+}
\tl{antaḥśūnyaṃ bahiḥśūnyaṃ śūnyakumbha ivāmbare ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_008_24]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*25}
\begin{translation}[hp04_008_25]
Do not think about the external or internal. [The yogi] should abandon all thought and think of nothing at all.
\end{translation}

%\begin{sources}[hp04_008_25]
%\end{sources}

\begin{testimonia}[hp04_008_25]
\emph{Haṃsavilāsa} p.~48
\begin{versinnote}
\tl{bāhyacintā na kartavyā tathaivāntaracintanam |\\+}
\tl{sarvacintāṃ parityajya na kiñcid api cintayet ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_008_25]
%°cintanā is not attested elsewhere.
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*26}
\begin{translation}[hp04_008_26]
The whole world is but a construct of mere ideation. A construct of mere ideation is an affectation of the mind. Jettison thought that is only ideation, take refuge in a resolve that is free of ideation, and obtain peace, o Rāma.
\end{translation}

\begin{sources}[hp04_008_26]
\emph{Laghuyogavāsiṣṭha} 7.27
\begin{versinnote}
\tl{saṃkalpajālakalanaiva jagat samagraṃ\\+}
\tl{saṃkalpajālakalanāt tu manovilāsaḥ |\\+}
\tl{saṃkalpamātram alam utsṛja nirvikalpam\\+}
\tl{āśritya niścayam avāpnuhi rāma śāntim ||\\!}
\end{versinnote}
\end{sources}

%\begin{testimonia}[hp04_008_26]
%\end{testimonia}

%\begin{philcomm}[hp04_008_26]
%\end{philcomm}

\begin{metre}[hp04_008_26]
Vasantatilakā
\end{metre}

%%%%%%%%%%
\subsection*{4.8*27}
\begin{translation}[hp04_008_27]
Just as camphor in fire and salt in water, so the mind, when being joined with the highest reality, dissolves.
\end{translation}

%\begin{sources}[hp04_008_27]
%\end{sources}

\begin{testimonia}[hp04_008_27]
\emph{Haṭharatnāvalī} 4.43
\begin{versinnote}
\tl{karpūram anale yadvat saindhavaṃ salile yathā |\\+}
\tl{tathā sandhīyamānaṃ hi manas tatraiva līyate||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_008_27]
%?? pāda c: hi better. Not in HP sources only HR. Adopt?
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*28}
\begin{translation}[hp04_008_28]
Mind is said to be all that is to be known, [all] that has been perceived, and [all] knowledge of that. Knowledge and what is to be known are destroyed together. There is no other path.
\end{translation}

%\begin{sources}[hp04_008_28]
%\end{sources}

\begin{testimonia}[hp04_008_28]
\emph{Haṭhatattvakaumudī} 51.35
\begin{versinnote}
\tl{jñeyaṃ sarvapratītaṃ ca tajjñānaṃ mana ucyate |\\+}
\tl{jñānaṃ jñeyaṃ samaṃ naṣṭaṃ nānyaḥ panthā dvitīyakaḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_008_28]
%Function of ca?
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*29}
\begin{translation}[hp04_008_29]
All this, everything moving and unmoving, is [just] a vision of the mind. When the mind has become free of the mind, it is called the absence of duality.
\end{translation}

\begin{sources}[hp04_008_29]
Gauḍapāda's \emph{Māṇḍūkyopaniṣatkārikā} 3.31
\begin{versinnote}
\tl{manodṛśyam idaṃ dvaitaṃ yat kiṃ cit sacarācaram |\\+}
\tl{manaso hy amanībhāve dvaitaṃ naivopalabhyate ||\\!}
\end{versinnote}

\emph{Amanaska} 2.79
\begin{versinnote}
\tl{manodṛśyam idaṃ sarvaṃ yat kim cit sacarācaram |\\+}
\tl{manaso hy unmanībhāve 'dvaitabhāvaṃ pracakṣate ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_008_29]
\emph{Yogacintāmaṇi} (attr.~to the \emph{Amanaska})
\begin{versinnote}
\tl{manodṛśyam idaṃ sarvaṃ yat kiṃ cit sacarācaram |\\+}
\tl{manasas tūnmanībhāvo 'dvaitābhāvaṃ pracakṣate ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_008_29]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*30}
\begin{translation}[hp04_008_30]
As a result of abandoning the things that are to be known the mind attains dissolution (\emph{layam}). When the mind has attained dissolution, isolation (\textit{kaivalyam}) remains.
\end{translation}

\begin{sources}[hp04_008_30]
\end{sources}

\begin{testimonia}[hp04_008_30]
\emph{Haṭharatnāvalī} 4.44
\begin{versinnote}
\tl{jñeyavastuparityāgād vilayaṃ yāti mānasaḥ |\\+}
\tl{mānase vilayaṃ yāte kaivalyam upajāyate ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_008_30]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*31}
\begin{translation}[hp04_008_31]
“Dissolution, dissolution”, they say. What kind of characteristics does dissolution have? Because subliminal impressions do not arise again, dissolution is the forgetting of the objects of the senses.
\end{translation}

%\begin{sources}[hp04_008_31]
%\end{sources}

\begin{testimonia}[hp04_008_31]
\emph{Haṭharatnāvalī} 1.13
\begin{versinnote}
\tl{layo laya iti prāhuḥ kīdṛśaṃ layalakṣaṇam |\\+}
\tl{apunarbhavasaṃsthānaṃ layo viṣayavismṛtiḥ ||\\!}
\end{versinnote}

\emph{Yogasārasaṅgraha} ??
\begin{versinnote}
\tl{śrīdattaḥ |\\+}
\tl{layo laya iti prāhur īdṛśaṃ lakṣaṇaṃ sphuṭam |\\+}
\tl{tatra sarvasamādhāne layo viṣayavismṛtiḥ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_008_31]
This verse may have been inspired by the \emph{Mokṣopāya} (e.g.~1.2.2).
\end{philcomm}

%%%%%%%%%%
\subsection*{4.8*32}
\begin{translation}[hp04_008_32]
Various methods like these, which are understood properly by one's own experience, are taught as paths to \textit{samādhi} by the great-souled teachers of former times.
\end{translation}

%\begin{sources}[hp04_008_32]
%\end{sources}

%\begin{testimonia}[hp04_008_32]
%\end{testimonia}

%\begin{philcomm}[hp04_008_32]
%\end{philcomm}

\begin{metre}[hp04_008_32]
Anuṣṭubh (c: bha-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.8*33}
\begin{translation}[hp04_008_33]
Homage to Suṣumṇā, to Kuṇḍalinī, to the nectar arising from the moon, to the mind beyond mind state, to you whose nature is consciousness, the great Śakti.%?? or whose nature is the consciousness of the great Śakti? JM: I have forgotten how we are writing nāḍīs etc. I like how we have everything here.
\end{translation}

%\begin{sources}[hp04_008_33]
%\end{sources}

%\begin{testimonia}[hp04_008_33]
%\end{testimonia}

%\begin{philcomm}[hp04_008_33]
%\end{philcomm}

\begin{metre}[hp04_008_33]
Anuṣṭubh (a: ra-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.8*34}
\begin{translation}[hp04_008_34]
The cultivation of the inner sound taught by Gorakṣa [and] approved even for deluded people for whom realisation of the highest reality is impossible is now taught.
\end{translation}

%\begin{sources}[hp04_008_34]
%\end{sources}

%\begin{testimonia}[hp04_008_34]
%\end{testimonia}

%\begin{philcomm}[hp04_008_34]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.9}
\begin{translation}[hp04_009]
The twelve and a half million methods of dissolution taught by  glorious Śiva are victorious. We think only the singular concentration on the inner sound is the most venerable of [all] the dissolutions.%??JM: How about "Glory to..." instead of "are victorious"? And then ""We consider one of the dissolutions in particular to be especially worthy of honour, concentration on the inner sound."
\end{translation}

\begin{sources}[hp04_009]
\emph{Yogatārāvalī} 2
\begin{versinnote}
\tl{sadā śivoktāni sapādalakṣa-\\+}
\tl{layāvadhānāni lasantu loke |\\+}
\tl{nādānusandhānasamādhim ekaṃ\\+}
\tl{manyāmahe mānyatamaṃ layānām ||\\+}
\tl{\var{lasantu ] Sa : lasanti Pa : vasanti Ad W Eb Ec Ed : ca santi  Ea : hi sarva Ma}\\+}
\tl{\var{loke ]  Ma  W Eall : bhūmau  Ad Pa Sa}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_009]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{śrīādināthena sapādakoṭi-\\+}
\tl{layaprakārāḥ kathitā jayanti |\\+}
\tl{nādānusandhānakam eva kāryam\\+}
\tl{manyāmahe nānyatamaṃ layānām ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 1.12
\begin{versinnote}
\tl{śrīādināthena sapādakoṭi-\\+}
\tl{layaprakārāḥ kathitā jayantu  |\\+}
\tl{nādānusandhānakam eva kāryaṃ\\+}
\tl{manyāmahe mānyatamaṃ layānām ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_009]
%\end{philcomm}

\begin{metre}[hp04_009]
Upajāti
\end{metre}

%%%%%%%%%%
\subsection*{4.10}
\begin{translation}[hp04_010]
Seated in the pose of the liberated ones, the yogi should adopt \emph{śāṃbhavī mudrā} and, with his mind one-pointed, listen to the inner sound in his right ear.
\end{translation}

%\begin{sources}[hp04_010]
%\end{sources}

\begin{testimonia}[hp04_010]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{muktāsanasthito yogī mudrāṃ sandhāya śāṃbhavīm |\\+}
\tl{śṛṇuyād dakṣiṇe karṇe nādam antargataṃ sadā ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} (ms.~2244 f.~124r)
\begin{versinnote}
\tl{muktāsanasthito yogī mudrāṃ saṃdhāya śāṃbhavīṃ [|]\\+}
\tl{śṛṇuyād dakṣiṇe karṇe nādam ekāṃtike sudhīḥ  [||]\\+}
\tl{\var{dakṣiṇe karṇe ] B220,  dakṣirṇe 2244}\\+}
\tl{\var{ekāṃtike ] 2244, ekāṃtate B220}\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_010]
We read this verse here rather than at 4.50 because it follows on from the Śāmbhavī section and is unnecessary and repetitive at 4.50 because 4.51-52 explain the practice.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.11}
\begin{translation}[hp04_011]
A fire that has been set on wood disappears together with the wood; the mind set on the [inner] sound dissolves together with the [inner] sound.
\end{translation}

%\begin{sources}[hp04_011]
%\end{sources}

\begin{testimonia}[hp04_011]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{kāṣṭhe pravartito vahniḥ kāṣṭhena saha śāmyati |\\+}
\tl{nāde pravartitaṃ cittaṃ nādena saha līyate ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.15
\begin{versinnote}
\tl{kāṣṭhe pravartito vahniḥ kāṣṭhena saha līyate |\\+}
\tl{nāde pravartitaṃ cittaṃ nādena saha śāmyati ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} (ms. 2244 f.~124r)
\begin{versinnote}
\tl{kāṣṭhaiḥ pravartito vahniḥ kāṣṭhena saha śāmyati |\\+}
\tl{nāde pravartitaṃ cittaṃ nādena saha līyate ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_011]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.12}
\begin{translation}[hp04_012]
Having forgotten everything external, the mind becomes one with the inner sound like milk and water, then quickly dissolves into the space of consciousness.
\end{translation}

%\begin{sources}[hp04_012]
%\end{sources}

\begin{testimonia}[hp04_012]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{vismṛtya sakalaṃ bāhyaṃ nāde dugdhāmbuvan naraḥ |\\+}
\tl{ekībhūyātha sahasā cidākāśe vilīyate ||\\!}
\end{versinnote}

Upāsanāsārasaṅgraha (p. 106)
\begin{versinnote}
\tl{vismṛtya sakalaṃ bāhyaṃ nāde dagdhāṃbuvan manaḥ |\\+}
\tl{ekībhūtaṃ tanyā cittaṃ rājayogābhidānakaṃ ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} (ms. 2244 f.~124r)
\begin{versinnote}
\tl{vismṛtya sakalaṃ bāhyaṃ nāde dugdhāṃbuvan manaḥ |\\+}
\tl{ekībhūyātha sahasā cidākāśe vilīyate ||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 39
\begin{versinnote}
\tl{vismṛtya sakalaṃ bāhyaṃ nāde dugdhāmbuvan manaḥ |\\+}
\tl{ekībhūyātha sahasā cidākāśe vilīyate ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_012]
%\end{philcomm}

\begin{metre}[hp04_012]
Anuṣṭubh (c: na-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.13}
\begin{translation}[hp04_013]
Having become intent on indifference through regular practice, the ascetic should immediately concentrate on nothing but the inner sound, which brings about the [state] beyond mind.%??JM: sadyas is odd. Apparently it has an archaic meaning of "daily". Consider? Immediately doesn't make sense. Or take it with unmanīkaraṇam. And sadābhyāsena could go with avadhārayet.
\end{translation}

%\begin{sources}[hp04_013]
%\end{sources}

\begin{testimonia}[hp04_013]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{audāsīnyaparo bhūtvā sadābhyāsena saṃyamī |\\+}
\tl{unmanīkaraṇaṃ sadyo nādam evāvadhārayet ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.38
\begin{versinnote}
\tl{audāsīnyaparo bhūtvā sadābhyāsena saṃyamī |\\+}
\tl{unmanīkārakaṃ sadyo nādam eva sadābhyaset ||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 40
\begin{versinnote}
\tl{udāsīnas tato bhūtvā sadābhyāsena saṃyamī |\\+}
\tl{unmanīkārakaṃ sadyo nādam evāvadhārayet ||\\!}
\end{versinnote}

\emph{Upāsanāsārasaṅgraha} p.~106
\begin{versinnote}
\tl{audāsīnyaparo bhūtvā sadābhyāsena samyamī |\\+}
\tl{unmanīkaraṇam sadyo nādam evānu [...] yet ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_013]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.14}
\begin{translation}[hp04_014]
What kind of indifference?\\
In the cold season, [indifference towards] whether [one has] an open tent or a quilt, with regard to good nourishment, whether it is cow's milk or water, with regard to food, whether it is lots of alms or forest roots, and with regard to the vessel for food, whether it is the hand, bucket or beggar's bowl.
\end{translation}

%\begin{sources}[hp04_014]
%\end{sources}

\begin{testimonia}[hp04_014]
\emph{Yogacintāmaṇi} ref??
\begin{versinnote}
\tl{kīdṛśaṃ caudāsīnyam—\\+}
\tl{śīte kāle kāpaṭī vā paṭī vā
pathyāhāre gopayo vā payo vā |\\+}
\tl{bhakṣye bhikṣāvṛndam āraṇyakandam
pāṇau droṇī kā parā bhojapātram ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.7
\begin{versinnote}
\tl{śīte kāle dvau paṭī vā paṭī vā pathyāhāre gopayo vā payo vā |\\+}
\tl{bhakṣye bhojye vṛttim āraṇyakaṃ vā pāṇī droṇī ko 'pi vā bhakṣyapātre ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī}  54.39
\begin{versinnote}
\tl{audāsīnyaṃ śītakāle paṭī vā
pathyāhāro gopayo vā payo vā |\\+}
\tl{bhojyaṃ bhikṣāvṛndam āraṇyakandaṃ
pāṇī droṇī kāpi vā bhojyapātram ||\\!}
\end{versinnote}

\end{testimonia}

\begin{philcomm}[hp04_014]
On \emph{caupaṭī} in the first \emph{pāda}, McGregor (1994:s.v.) and Callewaert (2009:s.v.) in their dictionaries of modern and old Hindi both give the meaning "open all around" for \emph{caupaṭ}. McGregor derives it from Sanskrit \emph{catuṣpaṭṭa}. Molesworth (1857:s.v.) in his Marathi dictionary gives "A quadrangular expanse or space, esp. as open and extended: also a broad and level tract" for \emph{caupaṭṭā}
\end{philcomm}

\begin{metre}[hp04_014]
Śālinī 
\end{metre}

%%%%%%%%%%
\subsection*{4.15}
\begin{translation}[hp04_015]
Having forever abandoned all worry and all activity, as a result of focusing on nothing but the inner sound the mind dissolves into the inner sound.
\end{translation}

%\begin{sources}[hp04_015]
%\end{sources}

\begin{testimonia}[hp04_015]
\emph{Haṭharatnāvalī} 4.13
\begin{versinnote}
\tl{sarvacintāṃ samutsṛjya sarvaceṣṭāṃ ca sarvadā |\\+}
\tl{nādam evānusandhānān nāde cittaṃ vilīyate ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} (ms. 2244 f.~124r)
\begin{versinnote}
\tl{sarvaciṃtā[ṃ] parityajya sarvakāle ca sarvadā |\\+}
\tl{nādam evānusandhatte nāde cittaṃ vilīyate ||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 41
\begin{versinnote}
\tl{sarvacintāṃ samutsṛjya sarvaceṣṭāvivarjitaḥ |\\+}
\tl{nādam evānusaṃdadhyān nāde cittaṃ vilīyate  ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_015]
The occurrence of two finite verbs in the second hemistich of some witnesses (groups 2 and 3) is infelicitous and perhaps points to an error that arose from the \textit{nādam evānusandhānān} reading in the group one?? mss. It seems the original reading was probably \textit{nāda evānusandhānāt}, and the -m- crept in as a hiatus break.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.16}
\begin{translation}[hp04_016]
Beginning (\emph{ārambhaḥ}), activation (\emph{ghaṭaḥ}), accumulation (\emph{paricayaḥ}) and completion (\emph{niṣpattiḥ}): they are the stages of yoga in all yogas.
\end{translation}

\begin{sources}[hp04_016]
\emph{Amaraugha} 34
\begin{versinnote}
\tl{ārambhaś ca ghaṭaś caiva paricayaḥ prakīrtitaḥ |\\+}
\tl{niṣpattiḥ sarvayogeṣu cetasas tasya bhūmikāḥ ||\\+}
\tl{\var{paricaryaḥ prakīrtitaḥ Ga :
tathā paricaya samṛtaḥ Ae (\textit{unmetr.}) :
pracayaś ca tritīyakaḥ Ad Gb :
pracayaś ca tṛtīryakaḥ  Ba :
pracayaṃ ca tritīyakaṃ  Ta}\\+}
\tl{\var{cetasas tasya bhūmikāḥ ] $\beta$ : yogāvasthā caturvidhā $\theta$}\\!}%?? variants necessary?
\end{versinnote}

\emph{Amṛtasiddhi} 19.2
\begin{versinnote}
\tl{ārambhaś ca ghaṭaś caiva paricayas tṛtīyakaḥ |\\+}
\tl{niṣpannaḥ sarvaśeṣeṣu yogāvasthāḥ prakīrtitāḥ ||\\!}
\end{versinnote}

\emph{Śivasaṃhitā} 3.31
\begin{versinnote}
\tl{ārambhaś ca ghaṭaś caiva tathā paricayas tathā |\\+}
\tl{niṣpattiḥ sarvayogeṣu yogāvasthā bhavanti tāḥ ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_016]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{īśvaraprokte—\\+}
\tl{ārambhaś ca ghaṭaś caiva tathā paricayo 'pi ca |\\+}
\tl{niṣpattiḥ sarvayogeṣu yogāvasthā bhavanti tāḥ ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.17
\begin{versinnote}
\tl{ārambhaś ca ghaṭaś caiva tathā paricayas tathā |\\+}
\tl{niṣpattiḥ sarvayogeṣu yogāvasthā bhavanti tāḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_016]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.17}
\begin{translation}[hp04_017]
Now the beginning stage:\\
As a result of the piercing of the knot of Brahmā, bliss arises in the void [and] an unstruck sound which has various tones is heard in the body.
\end{translation}
%?? MD: better to adopt "tatra ārambhaḥ"? The first component is often introduced with "tatra" after the enumeration in the important mss.

\begin{sources}[hp04_017]
\emph{Amaraugha} 35
\begin{versinnote}
\tl{brahmagranthes tathā bhedād ānandaḥ śūnyasambhavaḥ |\\+}
\tl{vicitrakvaṇako dehe 'nāhataḥ śrūyate dhvaniḥ ||\\+}
\tl{\var{°granthes tathā ] Ga : °granthes tato Ae : °granther bhaved Ba Gb Ta : °granthir bhaved Ad}\\+}
\tl{\var{bhedād ] Ga : bhedāt Ae : vedhād Ba Gb Ta : vedhā Ad}\\!}%?? variants needed?
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_017]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{brahmagranthir bhaved bhinna ānandaḥ śūnyasaṃbhavaḥ |\\+}
\tl{vicitrakṣaṇiko deho 'nāhataḥ śrūyate dhvaniḥ ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.18
\begin{versinnote}
\tl{brahmarandhre bhaved bhedo yo nādaḥ sūryasaṃbhavaḥ |\\+}
\tl{vicitrakvaṇado dehe 'nāhataḥ śrūyate dhvaniḥ ||\\+}
\tl{\var{yo nādaḥ ] ānandaḥ \vl}\\+}
\tl{\var{vicitrakvaṇado ] vicitrakvaṇako \vl}\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.15
\begin{versinnote}
\tl{tatra ārambhaḥ –\\+}
\tl{brahmagranthir bhaved bhinnād ānandaḥ śūnyasambhavaḥ |\\+}
\tl{vicitrakvaṇiko dehe 'nāhataḥ śrūyate dhvaniḥ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_017]
Brahmānanda understands the voids associated with each stage to be places in the body.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.18}
\begin{translation}[hp04_018]
In the beginning [stage], which is [the first phase of] voidness and in which the heart is full [of bliss], one becomes a yogi with a divine body, splendour, a divine smell and freedom from disease.%?? space after śūnye in edition. ?? JM:  With a divine body, radiant, smelling heavenly, free from disease [and] his heart full [of bliss], in the void in the beginning [stage] the practitioner attains yoga.
\end{translation}

\begin{sources}[hp04_018]
\emph{Amaraugha} verse 36
\begin{versinnote}
\tl{divyadehaś ca tejasvī divyagandho hy arogavān |\\+}
\tl{saṃpūrṇahṛdaye śūnye tv ārambhe yogavān bhavet ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_018]
\emph{Haṭharatnāvalī} 4.19
\begin{versinnote}
\tl{divyadehaḥ sutejasvī divyagandhas tv arogavān |\\+}
\tl{saṃpūrṇahṛdaye śūnye tv ārambhe yogavān bhavet ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.18
\begin{versinnote}
\tl{tejasvī divyagandhaś ca divyadeho 'py arogavān |\\+}
\tl{sampūrṇahṛdaye śūnye tv ārambhe yogavān bhavet ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_018]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.19}
\begin{translation}[hp04_019]
Now the activation stage:\\
In the second stage the breath, after bringing about union, goes into the middle [channel]. Then the yogi has a firm posture [and] he becomes a gnostic and equal to a god.
\end{translation}

\begin{sources}[hp04_019]
\emph{Amaraugha} 37
\begin{versinnote}
\tl{dvitīye saṅghaṭīkṛtya vāyur bhavati madhyagaḥ |\\+}
\tl{dṛḍhāsano bhaved yogī jñānī devasamas tadā ||\\+}
\tl{\var{dvitīye saṅ- ] $\beta$ : dvitīyāyāṃ $\theta$}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_019]
\emph{Haṭharatnāvalī} 4.20
\begin{versinnote}
\tl{dvitīyāyāṃ ghaṭīkṛtya vāyur bhavati madhyagaḥ |\\+}
\tl{dṛḍhāsano bhaved yogī kāmadevasamas tadā ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.15
\begin{versinnote}
\tl{atha ghaṭāvasthā –\\+}
\tl{dvitīyāyāṃ ghaṭīṃ kṛtvā vāyur bhavati madhyagaḥ |\\+}
\tl{dṛḍhāsano bhaved yogī jñānī devasamas tathā ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_019]
The name of the second stage, \emph{ghaṭa} (and the related form \emph{ghaṭīkṛtya} found in the first \emph{pāda} of this verse) can be understood in at least three ways, as union, activation or pot (the latter with an alchemical connotation, on which see 3.12 and Mallinson and Szanto 2021:20–23). In the \emph{Dattātreyayogaśāstra} (verse 90) its primary meaning is union and the united pairs are \emph{prāṇa} and \emph{apāna}, \emph{jīvātman} and \emph{paramātman}, and \emph{nāda} and \emph{bindu}.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.20}
\begin{translation}[hp04_020]
Then, as a result of the piercing of the knot of Viṣṇu there is a hint of supreme bliss in total emptiness (\emph{atiśūnye}), and then the pounding sound of a \textit{bherī} drum occurs.%?? in ed join śabdas and tadā
\end{translation}

\begin{sources}[hp04_020]
\emph{Amaraugha} 38
\begin{versinnote}
\tl{viṣṇugranthes tato bhedāt paramānandasūcakaḥ |\\+}
\tl{atiśūnye vimardaś ca bherīśabdas tato bhavet ||\\+}
\tl{\var{atiśūnye ] Ad Gb Ta : atiśūnyo $\beta$ Ba}\\+}
\tl{\var{tato ] $\beta$ : tathā  Ba : tadā Ad Gb Ta}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_020]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{viṣṇugranthir yadā bhinnaḥ paramānandasūcakaḥ |\\+}
\tl{atiśūnyavibhedaś ca bherīśabdas tadā bhavet ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.21
\begin{versinnote}
\tl{viṣṇugranthes tathā bhedaḥ paramānandasūcakaḥ |\\+}
\tl{atiśūnye vimardaś ca bherīśabdas tathā bhavet ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.21
\begin{versinnote}
\tl{viṣṇugranthir yadā bhinnā paramānandasūcikā |\\+}
\tl{atiśūnyavibhedaś ca bherīśabdas tathā bhavet ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_020]
V3?? and HR have the reading of the \textit{Amaraugha} (\textit{atiśūnye vimardaś ca}), which was adapted from the Buddhist concept of a series of four void-like meditative states and moments. In the \emph{Amaraugha},  two of the moments were left out and two were reinterpreted (as in the case of \emph{vimarda}, which seems to have be retained with the more general sense of a pounding sound). %It is possible that Svātmārāma adopted this reading but he may have also changed it to \textit{atiśūnyavibheda} (as attested by groups 2 and 3)  in an attempt to make sense of the pāda (i.e., the variety, atiśūnya, arises).??JM: I percented this as we've gone with alpha1. In fact do we need a note here? How about a more general note on the stages at the beginning of this section, referencing Amṛtasiddhi, Amaraugha and DYŚ?
See Birch 2019: 968–969; Mallinson and Szanto 2021:19–20.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.21}
\begin{translation}[hp04_021]
Now the accumulation stage:\\
In the third [stage], having pierced [the knot of Viṣṇu], the sound of a \textit{mardala} drum arises in the sky. Then [the yogi] reaches the great void, the abode of all supernatural powers.%?? vihāyas = sky, not void so I've changed tr.
\end{translation}
%

\begin{sources}[hp04_021]
\emph{Amaraugha} 39
\begin{versinnote}
\tl{tṛtīyāyāṃ tato bhittvā ninādo mardaladhvaniḥ |\\+}
\tl{mahāśūnyaṃ tato jātaṃ sarvasiddhisamāśrayam ||\\+}
\tl{\var{bhittvā ninādo ] Ga : bhītvādinādau Ae : bhittvā vipāko Ba Ad Gb Ta }\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_021]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{tṛtīyāyāṃ tato bhitvā vimāyo mardaladhvaniḥ |\\+}
\tl{mahāśūnyaṃ tathā yāti sarvasiddhisamāśrayam ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.22
\begin{versinnote}
\tl{tṛtīyāyāṃ tato nityaṃ āviṣkāro marddladhvaniḥ |\\+}
\tl{mahāśūnyaṃ tato yāti sarvasiddhisamāśrayaḥ ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.22
\begin{versinnote}
\tl{atha paricayāvasthā ||\\+}
\tl{tṛtīyāyāṃ tato jitvā sahajānandasambhavaḥ |\\+}
\tl{doṣaduḥkhajarāmṛtyuḥ kṣudhānidrāvivarjitaḥ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_021]
Here the object of piercing is unspecified, but it is likely to refer to the knot of Viṣṇu mentioned in the previous verse.
%?? MD: Pāda c - tathā (Gr2) to be changed to tato (Gr3) or tadā (alpha*,V3,Jyo)? JM: tadā is good and tr. already translates that way
\end{philcomm}

%%%%%%%%%%
\subsection*{4.22}
\begin{translation}[hp04_022]
Now the perfection stage:\\
Having overcome the [supreme] bliss of the mind, there arises innate bliss. [The yogi becomes] free of disease, suffering, old age, death, hunger and sleep.
\end{translation}

\begin{sources}[hp04_022]
\emph{Amaraugha} 40
\begin{versinnote}
\tl{paramānandarocitvāt sahajānandasambhavaḥ |\\+}
\tl{doṣaduḥkhajarāmṛtyukṣudhānidrāvivarjitaḥ ||\\+}
\tl{\var{40a paramānanda° ] $\beta$; cittānandaṃ $\theta$ • °rocitvāt ] $\beta$; tato jitvā $\theta$}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_022]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{cittānandaṃ tato jitvā sahajānandasaṃbhavaḥ |\\+}
\tl{doṣaduḥkhakṣudhānidrājarāmṛtyuvivarjitaḥ ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.23
\begin{versinnote}
\tl{cidānandaṃ tato jitvā paramānandasaṃbhavaḥ |\\+}
\tl{doṣaduḥkhajarāmṛtyukṣudhānidrāvivarjitaḥ ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.22
\begin{versinnote}
\tl{tṛtīyāyāṃ tato jitvā sahajānandasambhavaḥ |\\+}
\tl{doṣaduḥkhajarāmṛtyuḥ kṣudhānidrāvivarjitaḥ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_022]
22a’s \emph{cittānāndaṃ tato jitvā} is perhaps a patch for the somewhat obscure \emph{paramānandarocitvāt} in the source text.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.23}
\begin{translation}[hp04_023]
Now the perfection stage:\\
Then, having pierced Rudra's knot, the breath goes to all the  [of deities in the body]. In the perfected [stage] the sound of a flute becomes the sound of a resonating lute.
\end{translation}

\begin{sources}[hp04_023]
\emph{Amaraugha} 41
\begin{versinnote}
\tl{rudragranthiṃ tato bhittvā sarvapīṭhagato 'nilaḥ |\\+}
\tl{niṣpanno vaiṇavaḥ śabdaḥ kvaṇadvīṇākvaṇo bhavet ||\\+}
\tl{\var{sarva° ] Ae : sarvaṃ Ga : sattva° $\theta$}\\+}
\tl{\var{niṣpanno ] Ga : niṣpannā Ae : niṣpattau Ad Gb T : ++ttau Ba }\\+}
\tl{\var{kvaṇadvīṇākvaṇo ] conj. : kvaṇañ cailakvaṇo S1 : kvaṇañ caiva kvaṇo S2 : kvaṇadvitakvaṇo L3 : kvaṇanvitakvaṇo L2 : kvaṇanvītakva+ L4 : kvaṇatbhakvaṇo L1}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_023]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{rudragranthiṃ tato bhitvā sarvapīṭhagato 'nilaḥ |\\+}
\tl{niṣṭhāto vaiṇavaḥ śabdaḥ kvaṇadvīṇākvaṇo bhavet ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.24
\begin{versinnote}
\tl{rudragranthiṃ tato bhitvā śarvapīṭhagato 'nilaḥ |\\+}
\tl{niṣpattau vaiṇavaḥ śabdaḥ kvanadvīṇākvaṇo bhavet ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.24
\begin{versinnote}
\tl{atha niṣpattiḥ –\\+}
\tl{rudragranthiṃ tato bhitvā śarvapīṭhagato'nalaḥ |\\+}
\tl{niṣpanno vaiṇavaḥ śabdo kvaṇadvīṇākvaṇo bhavet ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_023]
The reading we have adopted for the fourth \emph{pāda}, \emph{kvaṇadvīṇākvaṇo}, Birch’s conjecture in his edition of the \emph{Amaraugha} is supported by several testimonia and \emph{Amṛtasiddhi} 31.2, where it is said that the sound of a \emph{vīṇā} arises in the fourth stage.

The reading \emph{śarva}, i.e.~Śiva, found in some testimonia and the \emph{Jyotsnā} makes good sense, but in its description of this stage, the \emph{Amṛtasiddhi}, which is the ultimate source of this passage has \emph{sarva} (30.1).
\end{philcomm}

%%%%%%%%%%
\subsection*{4.24}
\begin{translation}[hp04_024]
Then the mind having become one [with the sound??] is called Rājayoga. That yogi becomes a creator and destroyer, an equal to god.%??see note
\end{translation}

\begin{sources}[hp04_024]
\emph{Amaraugha} 42
\begin{versinnote}
\tl{ekībhūtaṃ tadā cittaṃ rājayogābhidhānakam |\\+}
\tl{sṛṣṭisaṃhārakartāsau yogīśvarasamo bhavet ||+}
\tl{\var{rājayogābhidhānakam ] Ga Ba Ad Gb Ed ; rājayogo 'bhidhīyate Ae}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_024]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{ekībhūtaṃ tathā cittaṃ rājayogābhidhāyakam |\\+}
\tl{sṛṣṭisaṃhārakartāsau yogīśvarasamo bhavet ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_024]
The bahuvrīhi rājayogābhidhānakaṃ means that the mind is called rājayoga, which is very unusual. Perhaps °bhidhāyakaṃ is better, meaning "expresses rājayoga", but it is only in J7 and V3.

%??SS: Then the mind becomes one. This is called Rājayoga. JM: looks best to me

%??JM: yogīśvarasamo = equal to a yogi sovereign? Odd to take asau with yogī.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.24*1}
%\begin{translation}[hp04_024_1]
%\end{translation}

%\begin{sources}[hp04_024_1]
%\end{sources}

%\begin{testimonia}[hp04_024_1]
%\end{testimonia}

\begin{philcomm}[hp04_024_1]
See 4.32.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.25}
\begin{translation}[hp04_025]
Whether or not this is liberation, in this very state a great, unbroken pleasure, which is rich in the nectar of absorption, is attained from Rājayoga.
\end{translation}

%\begin{sources}[hp04_025]
%\end{sources}

\begin{testimonia}[hp04_025]
\emph{Yogacintāmaṇi} ??
\begin{versinnote}
\tl{astu vā māstu vā siddhir atraivākhaṇḍitaṃ sukham |\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.16
\begin{versinnote}
\tl{astu vā māstu vā muktir atraivākhaṇḍitaṃ mahat |\\+}
\tl{layāmṛtaṃ laye saukhyaṃ rājayogād avāpyate ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.35
\begin{versinnote}
\tl{astu vā māstu vā muktir atraivākhaṇḍitaṃ mahat |\\+}
\tl{layāmṛtamayaṃ saukhyaṃ rājayogād avāpyate ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_025]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.26}
\begin{translation}[hp04_026]
Rājayoga without Haṭha and Haṭha without Rājayoga do not succeed so [the yogi] should practise both until the perfection stage.
\end{translation}

\begin{sources}[hp04_026]
\emph{Śivasaṃhitā} 5.222
\begin{versinnote}
\tl{haṭhaṃ vinā rājayogo rājayogaṃ vinā haṭhaḥ |\\+}
\tl{na sidhyati tato yugmam āniṣpatteḥ samabhyaset |\\+}
\tl{tasmāt pravartate yogī haṭhe sadgurumārgataḥ ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_026]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{haṭhaṃ vinā rājayogaṃ rājayogaṃ vinā haṭham |\\+}
\tl{ye vai caranti tān manye prayāsaphalavarjitān iti ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 1.19
\begin{versinnote}
\tl{haṭhaṃ vinā rājayogo rājayogaṃ vinā haṭhaḥ |\\+}
\tl{vyāptiḥ syād avinābhūtā śrīrājahaṭhayogayoḥ ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 55.1
\begin{versinnote}
\tl{haṭhaṃ vinā rājayogo rājayogaṃ vinā haṭhaḥ  |\\+}
\tl{na sidhyati tato yugmam āniṣpatteḥ samabhyaset  ||\\!} % = HP 2.77
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_026]
%\end{philcomm}

\begin{metre}[hp04_026]
Anuṣṭubh (a: ra-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.27}
\begin{translation}[hp04_027]
I consider those who are ignorant of Rājayoga and engage in only Haṭha to be labourers who get no reward for their efforts.%??JM: farm labourers? ploughmen? It would be good to get the agricultural meaning in.
\end{translation}

%\begin{sources}[hp04_027]
%\end{sources}

%\begin{testimonia}[hp04_027]
%\end{testimonia}

%\begin{philcomm}[hp04_027]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.28}
\begin{translation}[hp04_028]
The supreme reality is the seed, Haṭha the ground and indifference water. With these three the creeper of good fortune, the no-mind state, immediately grows.
\end{translation}

%\begin{sources}[hp04_028]
%\end{sources}

\begin{testimonia}[hp04_028]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{nādo bījaṃ haṭhaḥ kṣetram audāsīnyaṃ jalaṃ smṛtam |\\+}
\tl{unmanīkalpalatikā sadya evodbhaviṣyati ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_028]
The meaning of \emph{tattva} here is unclear. It is a synonym for \emph{samādhi} and \emph{unmanī} in the next verses. Brahmānanda says that \emph{tattva} means \emph{citta}, which works well: the mind grows to the beyond-mind state.

The reading \emph{smṛtam} is also possible and well-attested (groups 2 and 3).%??JM: not in apparatus.

\end{philcomm}

\begin{metre}[hp04_028]
Anuṣṭubh (c: na-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.29}
\begin{translation}[hp04_029]
Rājayoga, samādhi, unmanī, manonmanī, amaraugha, advaita, nirālamba, nirañjana...%?? format. Italic? Capitals?
\end{translation}

%\begin{sources}[hp04_029]
%\end{sources}

\begin{testimonia}[hp04_029]
\emph{Yogacintāmaṇi} ??
\begin{versinnote}
\tl{haṭhapradīpikāyām—\\+}
\tl{rājayogaḥ samādhiś ca unmanī ca manonmanī |\\+}
\tl{amaraughaughacāndrīva nirālambaṃ nirañjanam ||\\!}
\end{versinnote}

\emph{Upāsanāsārasaṅgraha} p.~106
\begin{versinnote}
\tl{haṭhapradīpikāyāṃ |\\+}
\tl{rājayogas samādhiś cāpy unmanī ca manonmanī |\\+}
\tl{amaraugho pi cādvaitaṃ nirālambaṃ niraṃjanaṃ  ||\\!}
\end{versinnote}

\emph{Yogasārasaṅgraha} p.~60
\begin{versinnote}
\tl{rājayogaḥ samādhiś conmanī ca manonmanī |\\+}
\tl{śivayogo layas tatvaṃ śūnyāśūnyaṃ nirañjanam |\\!}
\end{versinnote}

\emph{Haṃsavilāsa} (p. 47)
\begin{versinnote}
\tl{rājayogaḥ samādhiśca unmanī ca manonmanī |\\+}
\tl{amaraughālayas tatra śūnyāc chūnya paraṃ padam ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_029]
%\emph{amaraughaughacāndrī} appears to be a mistake perhaps because the word \emph{amaraugha} became \emph{amarolī}, hence prompting the change to \emph{candrī}.
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.30}
\begin{translation}[hp04_030]
amanaska, laya, tattva, śūnyāśūnya, para pada, jīvanmukti, sahaja and turya are synonyms.
\end{translation}

\begin{sources}[hp04_030]
\emph{Yogacintāmaṇi} ??
\begin{versinnote}
\tl{amanasko layaś caiva śūnyāśūnyaṃ parāparam |\\+}
\tl{jīvanmuktiś ca sahajaṃ turyaṃ cety ekavācakam iti ||\\!}
\end{versinnote}

\emph{Upāsanāsārasaṅgraha} p.~106
\begin{versinnote}
\tl{ama[na]sko layas tatvaṃ śūnyāśūnyaparaṃ  padaṃ |\\+}
\tl{jīvanmuktiś ca sahajaṃ turyaṃ cety ekavācakaṃ  ||\\!}
\end{versinnote}

\emph{Yogasārasaṅgraha} p.~60
\begin{versinnote}
\tl{amanaskaṃ yathā caitan nirālambaṃ nirañjanam |\\+}
\tl{jīvanmuktiś ca sahajam ity adir hy ekavācakam ||\\!}
\end{versinnote}
\end{sources}

%\begin{testimonia}[hp04_030]
%\end{testimonia}

%\begin{philcomm}[hp04_030]
%tattva may have dropped out of Gr2 and 3 because of the awkwardness of including it after tattvaṃ bījam in the previous verse (if tattva was considered something else).
%\end{philcomm}

\begin{metre}[hp04_030]
Anuṣṭubh (c: na-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.31}
\begin{translation}[hp04_031]
Two paths for the quick attainment of Unmanī are approved by me: tattva, the supreme bliss, and honouring the inner sound...
\end{translation}

%\begin{sources}[hp04_031]
%\end{sources}

\begin{testimonia}[hp04_031]
\emph{Yogacintāmaṇi} ??
\begin{versinnote}
\tl{unmanyavāptaye śīghraṃ dvau mārgau mama saṃmatau |\\+}
\tl{tatvaṃ paramasaukhyaṃ vā nādopāsanam eva vā ||\\!}
\end{versinnote}

\emph{Upāsanāsārasaṅgraha} p.~106
\begin{versinnote}
\tl{unmanyavāptaye śīghraṃ mārgau dvau mama sammatau |\\+}
\tl{tatvaṃ paramasākhyaṃ vā nādopāsanam eva vā  ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_031]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.32}
\begin{translation}[hp04_032]
which is approved even for stupid people whose minds are immersed in [worldly] pleasures. Dissolution which instantly bestows bliss arises from the inner sound.
\end{translation}

%\begin{sources}[hp04_032]
%\end{sources}

\begin{testimonia}[hp04_032]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{saukhyapraviṣṭacittānāṃ mūḍhānām api saṃmatam |\\+}
\tl{sadya ānandasandhāyī jāyate nādajo layaḥ ||\\!}
\end{versinnote}

\emph{Upāsanāsārasaṅgraha} p.~106
\begin{versinnote}
\tl{sāṅkhye praviṣṭacittānāṃ mūḍhānām api saṃmateḥ |\\+}
\tl{tasya svānaṃdasa [...] ryo jāyate nādajo layaḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_032]
%J5 has saṃmataḥ, agreeing with layaḥ
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.32*1}
\begin{translation}[hp04_032_1]
There is one seed [syllable] consisting of creation; one \emph{mudrā}, \emph{khecarī}, one god, the unsupported, [and] one state, mind beyond the mind.
\end{translation}

\begin{sources}[hp04_032_1]
\emph{Tantrāloka} 32.64
\begin{versinnote}
\tl{ekaṃ sṛṣṭimayaṃ bījaṃ yadvīryaṃ sarvamantragam |\\+}
\tl{ekā mudrā khecarī ca mudraughaḥ prāṇito yayā ||\\!}
\end{versinnote}

\emph{Tantrālokaviveka} 32.63
\begin{versinnote}
\tl{yad āgamaḥ –\\+}
\tl{ekaṃ sṛṣṭimayaṃ bījam ekā mudrā ca khecarī	|\\+}
\tl{dvāvekaṃ yo vijānāti sa vai pūjyaḥ kulāgame	||\\!}
\end{versinnote}
\emph{Śivasūtravimarśinī} 5
\begin{versinnote}
\tl{ekaṃ sṛṣṭimayaṃ [sṛṣṭimayaṃ bījam iti mantravīryarūpam aham iti bījam | 
mudrā parabhairavīyātmā |] bījam ekā mudrā ca khecarī |\\+}
\tl{dvāv etau yasya jāyete so'tiśāntapade sthitaḥ ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_032_1]
\emph{Haṭharatnāvalī} 4.28
\begin{versinnote}
\tl{ekaṃ sṛṣṭimayaṃ bījam ekā mudrā ca khecarī |\\+}
\tl{eko devo nirālambaḥ ekāvasthā manonmanī||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣad} 52cd-53ab
\begin{versinnote}
\tl{śaṅkhadundubhinādaṃ ca na śruṇoti kadācana || 52 ||\\+}
\tl{kāṣṭhavajjñāyate deha unmanyāvasthayā dhruvam |\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_032_1]
See 3.48.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.32*2}
\begin{translation}[hp04_032_2]
[The yogi] never hears the sounds of [even] conch shells and large drums. As a result of the state of no mind, the body assuredly becomes as [insentient as a piece of] wood.
\end{translation}

\begin{sources}[hp04_032_2]
Jñānasāra 3.7
\begin{versinnote}
\tl{śaṅkhadundubhinādena na śṛṇoti kadācana |\\+}
\tl{kāṣṭava[j] jñāyate yogī notpattyā vai prajāyate ||\\!}
\end{versinnote}
\end{sources}

%\begin{testimonia}[hp04_032_2]
%\end{testimonia}

%\begin{philcomm}[hp04_032_2]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.32*3}
\begin{translation}[hp04_032_3]
Free from all states of mind and all thought, the yogi is as if dead. He is liberated. In this there is no doubt.
\end{translation}

%\begin{sources}[hp04_032_3]
%\end{sources}

\begin{testimonia}[hp04_032_3]
\emph{Nādabindūpaniṣad} 51cd-52ab
\begin{versinnote}
\tl{sarvāvasthāvinirmuktaḥ sarvacintāvivarjitaḥ || 51 ||\\+}
\tl{mṛtavat tiṣṭhate yogī sa mukto nātra saṃśayaḥ |\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 51.75
\begin{versinnote}
\tl{sarvāvasthāvinirmuktaḥ sarvacintāvivarjitaḥ |\\+}
\tl{mṛtavattiṣṭhate yogī sa mukto nātra saṃśayaḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_032_3]
%Maybe kāṣṭhavat crept in as a repetition of the previous verse.
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.32*4}
\begin{translation}[hp04_032_4]
The yogi in \textit{samādhi} experiences neither cold nor heat, neither suffering nor pleasure, neither praise nor scorn.
\end{translation}

\begin{sources}[hp04_032_4]
\emph{Vivekamārtaṇḍa} 166
\begin{versinnote}
\tl{nābhijānāti śītoṣṇaṃ na duḥkhaṃ na sukhaṃ tathā |\\+}
\tl{na mānaṃ nāpamānaṃ ca yogī yuktaḥ samādhinā ||\\!}
\end{versinnote}
\end{sources}

%\begin{testimonia}[hp04_032_4]
%\end{testimonia}

%\begin{philcomm}[hp04_032_4]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.32*5}
\begin{translation}[hp04_032_5]
The yogi in \textit{samādhi} cannot be wounded by any weapon, killed by any person or overpowered by mantras and magic.
\end{translation}

\begin{sources}[hp04_032_5]
\emph{Vivekamārtaṇḍa} 168
\begin{versinnote}
\tl{abhedyaḥ sarvaśastrāṇām avadhyaḥ sarvadehinām |\\+}
\tl{agrāhyo mantratantrāṇāṃ yogī yuktaḥ samādhinā ||\\!}
\end{versinnote}
\end{sources}

%\begin{testimonia}[hp04_032_5]
%\end{testimonia}

%\begin{philcomm}[hp04_032_5]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.32*6}
\begin{translation}[hp04_032_6]
The yogi in \textit{samādhi} does not experience smell, taste, form, touch, sound, himself nor anyone else.
\end{translation}

\begin{sources}[hp04_032_6]
\emph{Vivekamārtaṇḍa} 165
\begin{versinnote}
\tl{na gandhaṃ na rasaṃ rūpaṃ na ca sparśaṃ na nisvanam |\\+}
\tl{nātmānaṃ na paraṃ vetti yogī yuktaḥ samādhinā ||\\!}
\end{versinnote}
\end{sources}

%\begin{testimonia}[hp04_032_6]
%\end{testimonia}

%\begin{philcomm}[hp04_032_6]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.32*7}
\begin{translation}[hp04_032_7]
He is indeed truly liberated whose mind is neither asleep nor awake, neither mindful nor otherwise, and neither stops nor starts.
\end{translation}

\begin{sources}[hp04_032_7]
\emph{Gorakṣaśataka} 7
\begin{versinnote}
\tl{cittaṃ na suptaṃ no jāgrat smṛtiman na ca nānyathā |\\+}
\tl{nāstam eti na codeti yasyāsau mukta eva sa ||\\!}
\end{versinnote}
\end{sources}

%\begin{testimonia}[hp04_032_7]
%\end{testimonia}

\begin{philcomm}[hp04_032_7]
The first line is corrupt in the \emph{Gorakṣaśataka} witnesses; its text has been constituted from the \emph{Haṭhapradīpikā}.
\end{philcomm}

\begin{metre}[hp04_032_7]
Anuṣṭubh (a: ma-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.32*8}
\begin{translation}[hp04_032_8]
[The yogi] who remains at ease as though asleep in the waking state, without breathing in and out, is definitely liberated.
\end{translation}

\begin{sources}[hp04_032_8]
\emph{Amanaska} 2.59
\begin{versinnote}
\tl{sadā jāgradavasthāyāṃ suptavad yo 'vatiṣṭhate |\\+}
\tl{niśvāsocchvāsahīnas ca niścitaṃ mukta eva saḥ ||\\+}
\tl{\var{sadā jāgradavasthāyāṃ ] \vl
sadā jāgṛvadasthāyāṃ,
sadā jāgrat apasthāyāṃ,
sadā jāgradavasther ya,
sa jāgras tadavasthāyāṃ,
yadā jāgṛdavasthāyāṃ,
yo jāgrad yad avasthāyāṃ,
svapnajāgradavasthāyāṃ,
suptajāgradavasthāyāṃ}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_032_8]
\emph{Kulārṇavatantra} 9.11
\begin{versinnote}
\tl{svapnajāgradavasthāyāṃ suptavat yo 'avatiṣṭhate | \\+}
\tl{niśvāsocchvāsahīnaśca niścitaṃ mukta eva saḥ || 11 ||\\!}
\end{versinnote}

\emph{Yuktabhavadeva} 1.64
\begin{versinnote}
\tl{svapnajāgradavasthāyāṃ suptavat yo 'vatiṣṭhate |\\+}
\tl{niśvāsocchvāsahīnaś ca niścitaṃ mukta eva saḥ ||64||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 55.24
\begin{versinnote}
\tl{sadā jāgradavasthāyāṃ suptavad yo ‘vatiṣṭhate |\\+}
\tl{niśvāsocchvāsavihīnaś ca niścitaṃ mukta eva saḥ || 55.24 = AY 2.59\\!}
\end{versinnote}

\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{sadā jāgradavasthāyāṃ suptavada yo'vatiṣṭhate ||\\+}
\tl{niḥśvāsocchāsahīnaśca niścitaṃ mukta eva saḥ|\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_032_8]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.33}
\begin{translation}[hp04_033]
Only the glorious guru lord knows the unique ineffable bliss that has arisen in the hearts of yogi lords who experience \textit{samādhi} by concentrating on the inner sound.
\end{translation}

\begin{sources}[hp04_033]
Yogatārāvalī 3
\begin{versinnote}
\tl{nādānusandhānasamādhibhājāṃ yogīśvarāṇāṃ hṛdaye pragūḍham |\\+}
\tl{ānandamātraṃ vacasām avācyaṃ jānāti taṃ śrīgurunātha ekaḥ ||\\+}
\tl{\var{ekaḥ ] Pa : eva Ad}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_033]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{nādānusandhānasamādhibhājāṃ
yogīśvarāṇāṃ hṛdaye prarūḍhaṃ |\\+}
\tl{ānandam ekaṃ vacasām avācyaṃ
jānāti tatvaṃ gurunātha eva ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.5
\begin{versinnote}
\tl{nādānusandhānasamādhibhājāṃ yogīśvarāṇāṃ hṛdaye prarūḍham |\\+}
\tl{ānandam ekaṃ vacaso 'py agamyaṃ jānāti taṃ śrīgurunātha eva ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_033]
%Check consistency on tr. of anusandhāna.
%\end{philcomm}

\begin{metre}[hp04_033]
Upajāti
\end{metre}

%%%%%%%%%%
\subsection*{4.33*1}
\begin{translation}[hp04_033_1]
Seated in the pose of the liberated ones, the yogi should adopt \emph{śāṃbhavī mudrā} and listen continuously to the inner sound in his right ear.
\end{translation}
%

%\begin{sources}[hp04_033_1]
%\end{sources}

\begin{testimonia}[hp04_033_1]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{muktāsanasthito yogī mudrāṃ sandhāya śāṃbhavīm |\\+}
\tl{śṛṇuyād dakṣiṇe karṇe nādam antargataṃ sadā ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} (ms. 2244 f.~124r)
\begin{versinnote}
\tl{muktāsanasthito yogī mudrāṃ saṃdhāya śāṃbhavīṃ [|]\\+}
\tl{śṛṇuyād dakṣiṇe karṇe nādam ekāṃtike sudhīḥ  [||]\\+}
\tl{\var{dakṣiṇe karṇe ] B220,  dakṣirṇe 2244}\\+}
\tl{\var{ekāṃtike ] 2244, ekāṃtate B220}\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 31
\begin{versinnote}
\tl{siddhāsane sthito yogī mudrāṃ sandhāya vaiṣṇavīm |\\+}
\tl{śṛṇuyād dakṣiṇe karṇe nādam antargataṃ sadā ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_033_1]
See 4.10.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.34}
\begin{translation}[hp04_034]
[The yogi] who desires yogic sovereignty should abandon all thought and concentrate with an attentive mind on nothing but the internal sound.
\end{translation}

%\begin{sources}[hp04_034]
%\end{sources}

\begin{testimonia}[hp04_034]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{sarvacintāṃ parityajya sāvadhānena cetasā |\\+}
\tl{nādam evānusandhatte yogasāmrājyam icchatā ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.14
\begin{versinnote}
\tl{sarvacintāṃ parityajya sāvadhānena cetasā |\\+}
\tl{nāda evānusandheyaḥ yogasāmrājyasiddhaye ||\\!}
\end{versinnote}

Haṭhasaṅketacandikā (f. 124r)
\begin{versinnote}
\tl{sarvacittaṃ parityajya sāvadhānena cetasā |\\+}
\tl{nāda evānusaṃdheyo yogasāmrājyam icchatā ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_034]
%N3 icchatā
%yogasāmrājya = rājayogasiddhi
%numbering is array in the apparatus
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.35}
\begin{translation}[hp04_035]
The sage should block his ears with cotton and fix the mind on the sound which he hears, until he attains a state of stillness.
\end{translation}

%\begin{sources}[hp04_035]
%\end{sources}

\begin{testimonia}[hp04_035]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{karṇau pidhāya hastena yaḥ śṛṇoti dhvaniṃ muniḥ |\\+}
\tl{tāvac cittaṃ sthiraṃ kuryād yāvat sthirapadaṃ vrajet ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.8
\begin{versinnote}
\tl{karṇau pidhāya tūlena yaḥ śṛṇoti dhvaniṃ yamī |\\+}
\tl{tatra cittaṃ sthiraṃ kuryād yāvat sthirapadaṃ vrajet ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} (f. 124r)
\begin{versinnote}
\tl{karṇau pidhāya hastābhyāṃ yaś śṛṇoti dhvaniṃ muniḥ |\\+}
\tl{tatra cittaṃ sthiraṃ kuryād yāvat sthirapadaṃ vrajet ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_035]
%adopt sthirīkuryād [MD: Done.]%??JM: remove space after sthirī  in ed.
%mūlena (alpha) probably came from tūlena. The hasta variations seem impractible because the yogi would have to hold the arms up, but may have been adopted by yogis who did not have access to cotton (?).??JM: I think these comments are unnecessary (and surely it is possible/practicable to use the fingers?).

Several witnesses have \emph{hastābhyām} or \emph{hastena} here. Blocking the ears with the fingers in order to listen to the inner sounds is attested as early as the \textit{Svacchandatantra}, in a practice called \textit{ṣaṇmukhīkaraṇa}, in which the other openings of the head are also blocked with the fingers (Vasudeva 2004: 272 n.66). \emph{Śivasaṃhitā} 5.36–46 teaches a similar practice.
%aṅguṣṭābhyām ubhe śrotre tarjanībhyāṃ vilocane
%nāsārandhre ca madhyābhyām anāmābhyāṃ mukhaṃ dṛḍham 5.36
%nirudhya mārutaṃ yogī yadaivaṃ kurute bhṛśam
%tadā tatkṣaṇamātmānaṃ jyotīrūpaṃ sa paśyati 5.37
%tattejo dṛśyate yena kṣaṇamātraṃ nirākulam
%sarvapāpavinirmuktaḥ sa yāti paramāṃ gatim 5.38
%nirantarakṛtābhyāsād yogī vigatakalmaṣaḥ
%sarvaṃ dehādi vismṛtya tadbhinnas tu svayaṃ bhavet 5.39
%yaḥ karoti sadābhyāsaṃ guptācāreṇa mānavaḥ
%sa vai brahmaṇi līnaḥ syāt pāpakarmarato yadi 5.40
%gopanīyaḥ prayatnena sadyaḥpratyayakārakaḥ
%nirvāṇadāyako loke yogo 'yaṃ mama vallabhaḥ 5.41
%nādaḥ saṃjāyate tasya krameṇābhyāsataś ca vai
%mattabhṛṅgāvalīvīṇāsadṛśaḥ prathamo dhvaniḥ 5.42
%evam abhyāsataḥ paścāt saṃsāradhvāntanāśanaḥ
%ghaṇṭāravasamaḥ paścād dhvanir megharavopamaḥ 5.43
%dhvanau tasmin mano dattvā yadā tiṣṭhati nirbharam
%tadā saṃjāyate siddhir layasya mama vallabhā 5.44
%tatra nāde yadā cittaṃ ramate yogino bhṛśam
%vismṛtya sakalaṃ bāhyaṃ nādena saha śāmyati 5.45
%etadabhyāsayogena jitvā samyag guṇān bahūn
%sarvārambhaparityāgī cidākāśe vilīyate 5.46
In the \textit{Haṭhayogasaṃhitā} (p. 68), the practice of blocking the ears with the hands is stipulated for \textit{bhrāmarī kumbhaka}.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.36}
\begin{translation}[hp04_036]
When this inner sound is being cultivated, it drowns out external sound. After a fortnight the yogi overcomes all distraction and becomes happy.
\end{translation}

%\begin{sources}[hp04_036]
%\end{sources}

\begin{testimonia}[hp04_036]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{abhyasyamāno nādo 'yaṃ bāhyam āvartayed dhvanim |\\+}
\tl{paścād vikṣepam akhilaṃ jitvā yogī sukhī bhavet ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} (f. 124r)
\begin{versinnote}
\tl{abhyasyamāno nādo yaṃ bāhyam āvarttayet dhvaniṃ |\\+}
\tl{pakṣād vikṣepam akhilaṃ jitvā yogī sukhībhavet ||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 32
\begin{versinnote}
\tl{abhyasyamāno nādo 'yaṃ bāhyam āvṛṇute dhvanim  |\\+}
\tl{pakṣād vipakṣam akhilaṃ jitvā turyapadaṃ vrajet ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_036]
%\end{philcomm}

\begin{metre}[hp04_036]
Anuṣṭubh (a: ma-vipulā; c: na-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.37}
\begin{translation}[hp04_037]
In the first stage of practice, a loud sound of various kinds is heard. Then, as the practice progresses, a quieter and quieter sound is heard.
\end{translation}

%\begin{sources}[hp04_037]
%\end{sources}

\begin{testimonia}[hp04_037]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{śrūyate prathamābhyāse nādo nānāvidho mahān |\\+}
\tl{vartamāne tato 'bhyāse śrūyate sūkṣmasūkṣmataḥ ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.9
\begin{versinnote}
\tl{śrūyate prathamābhyāse nādo nānāvidho bahuḥ |\\+}
\tl{vardhamāne tato 'bhyāse śrūyate sūkṣmasūkṣmataḥ ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.31
\begin{versinnote}
\tl{śrūyate prathamābhyāse nādo nānāvidho bahuḥ |\\+}
\tl{vardhamāne tato 'bhyāse śrūyate sūkṣmasūkṣmataḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_037]
%Adopt sūkṣmasūkṣmataḥ (alpha)[MD: Done.]
%sūkṣma seems to mean more than just quiet.
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.38}
\begin{translation}[hp04_038]
In the first stage, [the sounds] are those that are produced by the ocean, a [storm] cloud, a kettle drum and a waterfall. In the intermediate stage, they are [the sounds] produced by a bass drum and conch and a bell and trumpet.
\end{translation}

%\begin{sources}[hp04_038]
%\end{sources}

\begin{testimonia}[hp04_038]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{ādau jaladhijīmūtabherījharjharasaṃbhavaḥ |\\+}
\tl{madhye mardalaśaṅkhottha ghaṇṭākāhalakās tathā ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.10
\begin{versinnote}
\tl{ādau jaladhijīmūtabherīnirjharasaṃbhavāḥ |\\+}
\tl{madhye marddalaśaṃkhotthā ghaṇṭākāhalakās tathā ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.32
\begin{versinnote}
\tl{ādau jaladhijīmūtabherīnirjharasambhavaḥ |\\+}
\tl{madhye marddalaśaṃkhotthā ghaṃṭākāhalakās tathā ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_038]
?? more refs there are different options for kāhala, such as a trumpet or drum that looks like datura flower. See Schmidt, Nachträge: "eine Art Trommel (kāhalāsu dhattūrapuṣpākāramukhabherīṣu), Yaśastilakam (commentary).
\end{philcomm}

%%%%%%%%%%
\subsection*{4.39}
\begin{translation}[hp04_039]
In the final stage, there are the sounds of little bells, a bamboo flute, a veena and a bee.

[These] various sound are heard in the body.
\end{translation}

%\begin{sources}[hp04_039]
%\end{sources}

\begin{testimonia}[hp04_039]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{anye tu kiṅkiṇīvṛndavīṇābhramaraniḥsvanāḥ |\\+}
\tl{iti nānāvidho nādaḥ śrūyate dehamadhyagaḥ ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.11
\begin{versinnote}
\tl{ante tu kiṃkiṇīvṛndavīṇābhramaraniḥsvanāḥ |\\+}
\tl{iti nānāvidhā nādāḥ śrūyante dehamadhyataḥ ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.33
\begin{versinnote}
\tl{ante tu kiṃkiṇī vaṃśanādā bhramaraniḥsvanāḥ |\\+}
\tl{iti nānāvidhā nādāḥ śrūyante yatra madhyataḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_039]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.40}
\begin{translation}[hp04_040]
Even if a loud noise such as that of a [storm] cloud or kettle drum is being heard, the [yogi] should concentrate on only the very quietest sound in it.
\end{translation}

%\begin{sources}[hp04_040]
%\end{sources}

\begin{testimonia}[hp04_040]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{mahati śrūyamāṇe 'pi meghabheryādike dhvanau |\\+}
\tl{tataḥ sūkṣmāt sūkṣmataraṃ nādam eva parāmṛśet ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.34
\begin{versinnote}
\tl{mahati śrūyamāṇe 'pi meghabheryādike svane |\\+}
\tl{tatra sūkṣmāt sūkṣmataraṃ nādam eva parāmṛśet ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_040]
%\end{philcomm}

\begin{metre}[hp04_040]
Anuṣṭubh (c: bha-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.41}
\begin{translation}[hp04_041]
Or, the mind should filter out the gross sound for the subtle, or the subtle for the gross, or, abandoning both, be [focused] on [a sound in] the middle, [and the] should not move it elsewhere.%??JM: here gross and subtle do seem to make more sense
\end{translation}

%\begin{sources}[hp04_041]
%\end{sources}

\begin{testimonia}[hp04_041]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{ghanam utsṛjya vā sūkṣmaṃ sūkṣmaṃ pramṛjya vā ghanam |\\+}
\tl{paraṃ tatraiva niḥkṣipya mano nānyatra cālayet ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.35
\begin{versinnote}
\tl{ghanam utsṛjya vā sūkṣme sūkṣmam utsṛjya vā ghane |\\+}
\tl{ramamāṇam api kṣipraṃ mano nātra pracālayet ||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 37
\begin{versinnote}
\tl{ghanam utsṛjya vā sūkṣme sūkṣmam utsṛjya vā ghane |\\+}
\tl{ramamāṇam api kṣiptaṃ mano nānyatra cālayet ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_041]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.42}
\begin{translation}[hp04_042]
Alternatively, the mind fixes upon whatever sound it first attaches to and dissolves together with it.
\end{translation}

%\begin{sources}[hp04_042]
%\end{sources}

\begin{testimonia}[hp04_042]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{yatra kutrāpi vā nāde prathamaṃ viśate manaḥ |\\+}
\tl{tatraiva susthiraṃ kuryāt tena sārdhaṃ vilīyate ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.36
\begin{versinnote}
\tl{yatra kutrāpi vā nāde lagati prathamaṃ manaḥ |\\+}
\tl{tatraiva susthirībhūtvā tena sārdhaṃ vilīyate ||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 37
\begin{versinnote}
\tl{yatra kutrāpi vā nāde lagati prathamaṃ manaḥ  |\\+}
\tl{tatra tatra sthirībhūtvā tena sārdhaṃ vilīyate ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_042]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.43}
\begin{translation}[hp04_043]
Just as a bee drinking nectar has no regard for fragrances, so the mind attached to the inner sound does not desire the objects of the senses.
\end{translation}

%\begin{sources}[hp04_043]
%\end{sources}

\begin{testimonia}[hp04_043]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{makarandaṃ piban bhṛṅgo gandhān nāpekṣate yathā |\\+}
\tl{nādāsaktaṃ tathā cittaṃ viṣayān na hi kāṅkṣati ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.12
\begin{versinnote}
\tl{makarandaṃ pibed bhṛṅgo gandho na prekṣyate yathā |\\+}
\tl{nādāsaktaṃ tathā cittaṃ viṣayān na hi kāṃkṣate ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.41
\begin{versinnote}
\tl{makarandaṃ piban bhṛṃgo gandhaṃ nāpekṣate yathā |\\+}
\tl{nādāsaktaṃ tathā cittaṃ viśayān naiva kāṃkṣati ||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 42
\begin{versinnote}
\tl{makarandaṃ piban bhṛṅgo gandhān nāpekṣate tathā |\\+}
\tl{nādāsaktaṃ sadā cittaṃ viṣayaṃ na hi kāṅkṣati ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_043]
%gandhān is the alpha reading and is possible. [MD: adopted]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.44}
\begin{translation}[hp04_044]
When the mercury of the mind is bound and has cast off its fickle nature because it has been assimilated with the sulphur of the internal resonance, it attains the immobility called the unsupported [\textit{samādhi}].
\end{translation}

%\begin{sources}[hp04_044]
%\end{sources}

\begin{testimonia}[hp04_044]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{purā matsyendrabodhāya ādināthoditavacaḥ |\\+}
\tl{manaḥ pākam avāpnoti nirālambākhyaghoṭanam ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.42
\begin{versinnote}
\tl{baddhaṃ viyuktaṃ cāpalyaṃ nādagandhakajāraṇāt |\\+}
\tl{manaḥ pāradam āpnoti nirālambākhyakheṭakam ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_044]
See Hellwig 2009:204–206 on \emph{khoṭa}, “lame”, which in alchemy is a technical term use to describe mercury that has been processed many times using the \emph{māraṇa} technique and no longer moves.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.44*1}
\begin{translation}[hp04_044_1]
Bound by the sulphur of the inner sound, the lord that is the mercury of the mind immediately casts off its fickle nature and attains glory, known as “[the bird] with clipped wings”.
\end{translation}

%\begin{sources}[hp04_044_1]
%\end{sources}

\begin{testimonia}[hp04_044_1]
\end{testimonia}

\begin{philcomm}[hp04_044_1]

Cf.~\emph{Rasendracūḍāmaṇi} 16.52--55
\begin{versinnote}
\tl{evaṃ ca pañcamo grāsaḥ pradātavyo'ṣṭamāṃśataḥ |\\+}
\tl{sa pātrastho 'gnisaṃtapto na gacchati kathañcana ||16.52||\\+}
\tl{sa \textit{pakṣacchinna} ity uktaḥ sa mukto  'khiladurguṇaiḥ |\\+}
\tl{so'yaṃ niṣevitaḥ sūtas trimāsaṃ rājikāmitaḥ ||16.53||\\+}
\tl{viḍaṅgatriphalākṣaudraiḥ khe devaiḥ saha saṅgamam |\\+}
\tl{ghrāṇamātreṇa \textit{sūtendraḥ} sarvaroganikṛntanaḥ ||16.54||\\+}
\tl{guṇā ete vinirdiṣṭā rasasya rasavādibhiḥ |\\+}
\tl{sakalāste guṇāḥ satyā bhairaveṇa prakīrtitāḥ ||16.55||\\!}
\end{versinnote}

Cf.~also NWS (??ref) s.v.~\textit{pakṣaccheda}.

\end{philcomm}

\begin{metre}[hp04_044_1]
Anuṣṭubh (c: ma-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.45}
\begin{translation}[hp04_045]
As a result of listening to the inner sound, the  snake that is the mind forgets everything and, one-pointed, does not slip away anywhere.
\end{translation}

%\begin{sources}[hp04_045]
%\end{sources}

\begin{testimonia}[hp04_045]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{nādaśravaṇataś cittam antaraṅgaturaṅgamaḥ |\\+}
\tl{viśūnyaṃ sarvam ekāgryaṃ kutra cin na hi dhāvati ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.44
\begin{versinnote}
\tl{nādaśravaṇataś cittam antaraṃgakuraṃgakaḥ |\\+}
\tl{vismṛtya viśvam ekāgraḥ kutra cin na hi dhāvati ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_045]
The metaphor is that of snake charming: beguiled by the inner sound, the snake that is the mind becomes transfixed. Witnesses of the \gamma \ and \delta \ groups have \emph{turaṅgamaḥ} instead of \emph{bhujaṅgamaḥ}, perhaps because forms from \emph{dhāv} are unusual with the latter, but they are found at e.g.~\emph{Padmapurāṇa}, \emph{Garuḍapurāṇa} (refs ??). 
\end{philcomm}

%%%%%%%%%%
\subsection*{4.46}
\begin{translation}[hp04_046]
This inner sound is a sharpened goad with the power to restrain the bull elephant in must that is the mind as it wanders about in the garden of the sense objects.
\end{translation}

%\begin{sources}[hp04_046]
%\end{sources}

\begin{testimonia}[hp04_046]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{manomattagajendrasya viṣayodyānacāriṇaḥ |\\+}
\tl{niyāmanasamartho 'yaṃ ninādo niśitāṅkuśaḥ ||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 44cd–45ab
\begin{versinnote}
\tl{manomattagajendrasya viṣayodyānacāriṇaḥ ||\\+}
\tl{niyāmanasamartho 'yaṃ ninādo niśitāṅkuśaḥ |\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_046]
The unusual form \emph{niyāmana} is also found in Rasaśāstra works where it occurs in the context of restraining mercury and is a topic of discussion (e.g.~\emph{Rasaprakāśasudhākara} 1.23, \emph{Ānandakanda} 1.4.58–59).
\end{philcomm}

%%%%%%%%%%
\subsection*{4.47}
\begin{translation}[hp04_047]
The inner sound is a bolt for the swift horse of the mind so the yogi should regularly concentrate on cultivating it.
\end{translation}
% 4

%\begin{sources}[hp04_047]
%\end{sources}

\begin{testimonia}[hp04_047]
\emph{Haṭhatattvakaumudī} 54.46
\begin{versinnote}
\tl{antaraṃgaturaṃgasya vājinaḥ paridhāvataḥ |\\+}
\tl{nādopāstikhalīnaṃ hi niyāmanakaraṃ dṛḍham ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_047]
We understand \textit{ninādaḥ}, which is the subject of the previous line, as the subject in the first line of this verse. Several witness have \emph{paridhāyate} instead of \emph{parighāyate}. The latter is unattested but could mean "is a halter" based on \emph{abhidhānī}’s meaning of “halter”.
\end{philcomm}

\begin{metre}[hp04_047]
Anuṣṭubh (a: na-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.47*1}
\begin{translation}[hp04_047_1]
The inner sound is a net for trapping the deer of the mind and a hunter for trapping the antelope of the mind.
\end{translation}
% nādo’ntaraṅgasāraṅgabandhane vāgurāyate
%V15 antaraṅgaturaṅgasya nādo vyādhāyate pi ca

%\begin{sources}[hp04_047_1]
%\end{sources}

\begin{testimonia}[hp04_047_1]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{nādo 'ntaraṅgasāraṅgabandhane vāgurāyate ||\\+}
\tl{antaraṅgaturaṅgasya bandhane līyate 'pi ca ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 47
\begin{versinnote}
\tl{nādo'ntaraṃgasāraṃgabandhane vāgurāyate |\\+}
\tl{antaraṃgakuraṃgasya nādo vyādhāyate 'pi ca  ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_047_1]
Adopt V15 67cd

?? MD: rodhe may be correct. In a drive hunt, the animals are driven in a certain direction (often in a confined space)  by means of loud noises, etc., so that another hunter can easily shoot them. JM: yes, adopt, I have altered tr. to "trapping"
\end{philcomm}

%%%%%%%%%%
\subsection*{4.48}
\begin{translation}[hp04_048]
Striking the deer of the mind when it is transfixed by its attachment to inner sounds such as that of a bell, is very easy if the archer is skilful.
\end{translation}
% ghaṇṭādinādasaktastabdhāntaḥkaraṇahariṇasya |
% praharaṇam atisukaraṃ syāc charasaṃdhātā pravīṇaś cet ||
%

%\begin{sources}[hp04_048]
%\end{sources}

%\begin{testimonia}[hp04_048]
%\end{testimonia}

%\begin{philcomm}[hp04_048]
%N3 has a problem with the metre.
%\end{philcomm}

\begin{metre}[hp04_048]
Upagīti 
\end{metre}

%%%%%%%%%%
\subsection*{4.49}
\begin{translation}[hp04_049]
When the mind dissolves into that which is the most subtle object of perception in the unstruck sound, that is the supreme state of Viṣṇu.

The tone of that sound is that of the unstruck sound. The object of perception is inside the sound [and] the mind is inside the object of perception. When the mind dissolves [in it], that is the supreme state of Viṣṇu.
\end{translation}

% anāhatadhvaner antar jñeyaṃ yat sūkṣmasūkṣmakaṃ/
% manas tatra layaṃ yāti tad viṣṇoḥ paramaṃ padam//

\begin{sources}[hp04_049]
Uttaragītā 41cd–42 (source??)
\begin{versinnote}
\tl{anāhatasya śabdasya tasya śabdasya yo dhvaniḥ ||\\+}
\tl{dhvaner antargataṃ jyotir jyotirantargataṃ manaḥ |\\+}
\tl{tan mano vilayaṃ yāti tad viṣṇoḥ paramaṃ padam  ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_049]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{anāhatasya śabdasya tasya śabdasya yo dhvaniḥ ||\\+}
\tl{dhvaner antargataṃ jyotir jyotirantargataṃ manaḥ |\\+}
\tl{yan mano vilayaṃ yāti tad viṣṇoḥ paramaṃ padam ||\\!}
\end{versinnote}

Haṭhayogasaṃhitā p.~68
\begin{versinnote}
\tl{anāhatasya śabdasya tasya śabdasya yo dhvaniḥ |\\+}
\tl{dhvaner antargataṃ jyotir jyotiṣo 'ntargataṃ manaḥ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_049]
%?? to discuss. It appears that the following lines were added (in Gr2 and Gr3) as a note or supplement to 4.68. In particular, the hemistich anāhatasya śabdasya tasya... seems to comment on anāhatadhvaner.
\begin{versinnote}
\tl{anāhatasya śabdasya tasya śabdasya yo dhvaniḥ ||\\+}
\tl{dhvaner antargataṃ jyotir jyotirantargataṃ manaḥ |\\+}
\tl{tanmano vilayaṃ yāti tad viṣṇoḥ paramaṃ padam  || \\!}
\end{versinnote}
The source of them may be the Uttaragītā. However, in NGMPP E 2098-11 Uttaragītā ms these three lines are omitted from Kṛṣṇa's words, which start with oṃkāra.
%HI: tan could mean tatra.
%Looks like the last line was rewritten to include tatra.

%?? MD: I adopted the version transmitted in Gr4c and J10 (anāhatadhvaner antar jñeyaṃ yat sūkṣmasūkṣmakaṃ ...).
%?? Which is the more likely reading of the second line for the alternative version?
%A - dhvaner antargataṃ jyotir jyotirantargataṃ manaḥ (J5,G4,Gr2,Gr3)
%B - dhvaner antargataṃ jñeyaṃ jñeyasyāntargataṃ manaḥ (N3?,Jyo) JM: the second I think
\end{philcomm}

%%%%%%%%%%
\subsection*{4.50}
\begin{translation}[hp04_050]
As long as sound exists, there is a concept of space. That which is soundless is the supreme Brahman and is called the supreme self.
\end{translation}
% adopt samīryate J7, V3

%\begin{sources}[hp04_050]
%\end{sources}

\begin{testimonia}[hp04_050]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{tāvad ākāśasaṃkalpo yāvac chabdaḥ pravartate |\\+}
\tl{niḥśabdaṃ tatparaṃ brahma paramātmā samīryate ||\\!}
\end{versinnote}

Six-chapter \emph{Vivekamārtaṇḍa} 5.15
\begin{versinnote}
\tl{tāvad ākāśasaṅkalpo yāvac chabdaḥ pravartate |\\+}
\tl{niḥśabdaṃ tat parabrahma paramātmā sa gīyate || 15 ||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 47cd–48ab
\begin{versinnote}
\tl{tāvadākāśasaṅkalpo yāvacchabdaḥ pravatate |\\+}
\tl{niḥśabdaṃ tatparaṃ brahma paramātmā samīryate ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_050]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.51}
\begin{translation}[hp04_051]
Whatever is heard as the inner sound is nothing but Śakti. The formless one which hears it is nothing but the supreme lord.
\end{translation}
% adopt V3 reading nādarupeṇa

%\begin{sources}[hp04_051]
%\end{sources}

\begin{testimonia}[hp04_051]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{yat kiñ cin nāmarūpeṇa śrūyate śaktir eva sā |\\+}
\tl{yas tacchrotā nirākāraḥ sa eva parameśvaraḥ ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} f.~235r–235v
\begin{versinnote}
\tl{yat kiṃ cin nāmarūpeṇa śrūyate śaktir eva sā |\\+}
\tl{yasya śrottā nirākaraḥ sa eva parameśvaraḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_051]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.52}
\begin{translation}[hp04_052]
Blocking of the ears, mouth, eyes, and nose should not be performed [when] a clear inner sound is heard distinctly in the purified Suṣumṇā channel.%?? negative in first line? Adopt naiva?
\end{translation}

%\begin{sources}[hp04_052]
%\end{sources}

\begin{testimonia}[hp04_052]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{haṭhapradīpikāyām—\\+}
\tl{śravaṇapuṭanayananāsāpuṭarodhanaṃ kāryam |\\+}
\tl{śrīśuddhasuṣumṇāsaraṇau sphuṭam amalaḥ śrūyate nādaḥ ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} 123v–124r
\begin{versinnote}
\tl{haṭhapradīpikāyāṃ \\+}
\tl{śravaṇapuṭanayanayugulanāsāmukharodham eva kartavyaṃ |\\+}
\tl{śuddhasaṣumṇāśaraṇe sphuṭam amalaḥ śrūyate nādaḥ ||\\!}
\end{versinnote}

\emph{Saubhāgyalakṣmyupaniṣad} 4
\begin{versinnote}
\tl{śravaṇamukhanayananāsānirodhanenaiva kartavyam |\\+}
\tl{śuddhasuṣumnāsaraṇau sphuṭam amalaṃ śrūyate nādaḥ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_052]
%??N3 is missing an akṣara (nirodhaṃ aiva)
%??J5 has nirodhaṃ naiva
??Adopt alpha reading mukha°, naiva. HI: it is strongly attested (groups 4c, C6, etc) and more likely explains the change to °saṃrodhanaṃ kāryam
\end{philcomm}

\begin{metre}[hp04_052]
Upagīti 
\end{metre}

%%%%%%%%%%
\subsection*{4.52*1}
\begin{translation}[hp04_052_1]
Nāda is called Śakti. And gnosis of nāda is Sadāśiva. When gnosis of nāda has disappeared, only unmanī remains.
\end{translation}

%\begin{sources}[hp04_052_1]
%\end{sources}

\begin{testimonia}[hp04_052_1]
\emph{Haṭhatattvakaumudī} 54.50
\begin{versinnote}
\tl{nādaḥ śaktir iti khyāto nādajñānaṃ sadāśivaḥ |\\+}
\tl{nādajñāne vinaṣṭe ca tad unmany eva śiṣyate ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_052_1]
%??V15 reading for the third pāda is better than N19.
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.52*2}
\begin{translation}[hp04_052_2]
As long as there is the inner sound there is mind. At the end of the inner sound the mind beyond mind state [arises]. The void is said to be sonorous and Brahma is silent.
\end{translation}

%\begin{sources}[hp04_052_2]
%\end{sources}

\begin{testimonia}[hp04_052_2]
\emph{Haṭhatattvakaumudī} 54.51
\begin{versinnote}
\tl{nādo yāvan manas tāvan nādānte ca manonmanī |\\+}
\tl{saśabdaṃ kathitaṃ vāte niḥśabdaṃ brahma kathyate ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_052_2]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.52*3}
\begin{translation}[hp04_052_3]
When the accumulated subliminal impressions have been destroyed as a result of the continuous fusing of the mind with the internal resonance, the mind and breath are sure to dissolve into the unadorned [god] (\emph{nirañjane}).
\end{translation}

%\begin{sources}[hp04_052_3]
%\end{sources}

\begin{testimonia}[hp04_052_3]
\emph{Upāsanāsārasaṅgraha} ??
\begin{versinnote}
\tl{sadā nādānusandhānāt saṃkṣīṇe vāsanā bhavet |\\+}
\tl{nirañjane vilīneti niścalitaṃ mārutaṃ manaḥ ||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 49
\begin{versinnote}
\tl{saśabdaś cākṣare kṣīṇe niḥśabdaṃ paramaṃ padam |\\+}
\tl{sadā nādānusandhānāt saṃkṣīṇā vāsanā bhavet ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.52
\begin{versinnote}
\tl{sadā nādānusandhānāt saṃkṣīṇe vāsanākṣaye |\\+}
\tl{nirañjane ca līyate niścitaṃ cittamārutau ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_052_3]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.52*4}
\begin{translation}[hp04_052_4]
Thousands of crores of inner sounds and hundreds of crores of visual focal points all dissolve into the place where the unadorned god (\emph{devo nirañjana}) is.

%There are thousands of Crores of resonances and thousands of crores of sonic matter (bindu). All dissolve into where the spotless god exists.
%?? tr. the spotless god, Nirañjana. ? and in previous
\end{translation}

%\begin{sources}[hp04_052_4]
%\end{sources}

\begin{testimonia}[hp04_052_4]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{nādakoṭisahasrāṇi bindukoṭiśatāni ca |\\+}
\tl{sarve tatra layaṃ yānti yatra devo nirañjana iti ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 54.53
\begin{versinnote}
\tl{nādakoṭisahasrāṇi bindukoṭiśatāni ca |\\+}
\tl{sarve tatra layaṃ yānti yatra devo nirañjanaḥ ||\\!}
\end{versinnote}

\emph{Nādabindūpaniṣat} 50cd--51ab
\begin{versinnote}
\tl{nādakoṭisahasrāṇi bindukoṭiśatāni ca ||\\+}
\tl{sarve tatra layaṃ yānti brahmapraṇavanādake |\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_052_4]
The expression \textit{devo nirañjanaḥ} (`unadorned god') refers to Śiva.

Compare the quotation in the Śabdakalpadruma, s.v.~\emph{dharmaghaṭa}:
\begin{quote}
“\dots ante yāti paraṃ sthānaṃ yatra devo nirañjanaḥ||"
iti bhaviṣyapurāṇoktā dharmaghaṭavratakathā samāptā ||
\end{quote}
\end{philcomm}

%%%%%%%%%%
\subsection*{4.52*5}
%\begin{translation}[hp04_052_5]
%\end{translation}

%\begin{sources}[hp04_052_5]
%\end{sources}

%\begin{testimonia}[hp04_052_5]
%\end{testimonia}

\begin{philcomm}[hp04_052_5]
See 4.74.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.53}
\begin{translation}[hp04_053]
Enough of the verbosity of a learned gathering! O friend, listen to this instruction formerly taught by Śiva for awakening Matsyendra.

\end{translation}

%\begin{sources}[hp04_053]
%\end{sources}

\begin{testimonia}[hp04_053]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{kāṣṭhagoṣṭhīprasaṅgena nādam antargataṃ śṛṇu |\\+}
\tl{purā matsyendrabodhāya ādināthoditaṃ vacaḥ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_053]
This verse is likely to be authorial, composed to introduce next two verses, which are from the \emph{Candrāvalokana}, a dialogue between Matsyendra and Śiva. 1.34 is another probable authorial verse and, like this one, it has the vocative \emph{sakhe}.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.54}
\begin{translation}[hp04_054]
As long as the moving breath does not enter the middle path, as long as bindu is not stable, restrained by the breath, as long as the ultimate truth (\emph{tattvam}) does not become as natural as the sky, then all that one says is deceitful, and false prattle.
\end{translation}

\begin{sources}[hp04_054]
\emph{Candrāvalokana} 14
\begin{versinnote}
\tl{yāvan naiva praviśati caran māruto madhyamārge\\+}
\tl{yāvad bindur na bhavati dṛḍhaḥ prāṇavātaprabandhaḥ |\\+}
\tl{yāvad vyomnā sadṛśa sarasaṃ jāyate nonmanatvaṃ {[}em.~vyomnā sadṛśam arasaṃ?]\\+}
\tl{tāvat sarvaṃ yadi ca vadate dambhamithyāpralāpaḥ ||\\+}
\tl{\var{14b prāṇavātaprabandhaḥ ] 4345 : prāṇaghātaprabuddhaḥ 75278 : prāṇavāyuḥ prabuddhaḥ 7970 : prāṇavātaprabuddhaḥ T00788
14d sarvam] 7970, T00788, 75278 : vighnaṃ 4345}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_054]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{haṭhapradīpikāyāṃ—\\+}
\tl{yāvan naiva praviśati caran māruto madhyamārge\\+}
\tl{yāvad bindur na bhavati dṛḍhaḥ prāṇavātaprabaddhaḥ |\\+}
\tl{yāvad vyomnaḥ sahajasadṛśaṃ jñāyate naiva tatvam \\+}
\tl{tāvat sarvaṃ vadati yad idaṃ dambhamithyāpralāpa iti\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 2.2
\begin{versinnote}
\tl{yāvan naiva praviśati caran māruto madhyamārgaṃ\\+}
\tl{yāvat sūkṣmo na bhavati dṛḍhaḥ prāṇavātaprabandhaḥ |\\+}
\tl{yāvad vyomnā sahajasadṛśaṃ jāyate naiva cittaṃ\\+}
\tl{tāvat sarvaṃ vadati tad idaṃ dambhamithyāpralāpaḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_054]
%J5 has sthiraḥ (instead of dṛḍhaḥ) supporting the visarga (rather than dṛdhaṃ N3)
%tattva has been understood here (and in the verse with bīja) as the mind.
%\end{philcomm}

\begin{metre}[hp04_054]
Mandākrāntā 
\end{metre}

%%%%%%%%%%
\subsection*{4.55}
\begin{translation}[hp04_055]
Having learnt the correct piercing of Suṣumṇā, [the yogi] should make the breath go into the central channel, take it to the place of the moon and stop [the flow] in the nostrils.
\end{translation}
%nītvāsāv aindavasthāne

\begin{sources}[hp04_055]
\emph{Candrāvalokana} 32
\begin{versinnote}
\tl{jñātvā suṣumnāsadbhedaṃ kṛtvā vāyuṃ ca madhyagam |\\+}
\tl{kṛtvāsāv aindave sthāne ghrāṇarandhre nirodhayet ||\\+}
\tl{\var{satbhedaṃ ] 4345, 4340, T00788 : tatbhedaṃ  7970.}\\+}
\tl{\var{kṛtvāsāv aindave ] 7970, 4340, T00788 : kṛtvā*d*baindave 4345}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_055]
\emph{Upāsanāsārasaṅgraha} p.~31 (attributed to the \emph{Candrāvalokana})
\begin{versinnote}
\tl{jñatvā suṣumnāṃ tadbhedaṃ kṛtvā vāyuṃ ca madhyagam |\\+}
\tl{kṛtvāsau baindavasthāne ghraṇarandhre nirodhayet ||\\!}
\end{versinnote}

\emph{Yogakuṇḍalinyupaniṣat} 7cd–8ab
\begin{versinnote}
\tl{jñātvā suṣumnāṃ tadbhedaṃ kṛtvā vāyuṃ ca madhyagam ||\\+}
\tl{sthitvāsau baindavasthāne ghrāṇarandhre nirodhayet |\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} f.~107v–108r
\begin{versinnote}
\tl{jñātvā suṣumnāsadbhedaṃ kṛtvā vāyuṃ ca madhyagam |\\+}
\tl{sthitvā sadaiva svasthena prāṇarandhraṃ nirodhayet ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_055]
%??ghrāṇarandhra is too well supported to ignore, but the meaning is difficult to construe. JM: it's in alpha etc., why not adopt it? vāyum is object in pādas bcd.
%\end{philcomm}

\begin{metre}[hp04_055]
Anuṣṭubh (a: ma-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.56}
\begin{translation}[hp04_056]
The moon and sun move in iḍā and piṅgalā. The moon is said to be tamasic and the sun rajasic.
\end{translation}

\begin{sources}[hp04_056]
\emph{Vasiṣṭhasaṃhitā} 2.28–29ab
\begin{versinnote}
\tl{iḍāyāṃ piṅgalāyāṃ ca carataś candrabhāskarau |\\+}
\tl{iḍāyāṃ candramā jñeyaḥ piṅgalāyāṃ raviḥ smṛtaḥ ||\\+}
\tl{candras tāmasa ity uktaḥ sūryo rājasa ucyate |\\!}
\end{versinnote}

\emph{Yogayājñavalkya} 4.32cd–33
\begin{versinnote}
\tl{iḍāyāṃ piṅgalāyāṃ ca carataś candrabhāskarau ||\\+}
\tl{iḍāyāṃ candramā jñeyaḥ piṅglāyāṃ raviḥ smṛtaḥ |\\+}
\tl{candras tāmasa ity uktaḥ sūryo rājasa ucyate ||\\!}
\end{versinnote}

\emph{Matsyendrasaṃhitā} 4.41cd
\begin{versinnote}
\tl{iḍāyāṃ piṅgalāyāṃ ca parataś candrabhāskarau ||\\!} 
\end{versinnote}

\end{sources}

\begin{testimonia}[hp04_056]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{iḍāyāṃ piṃgalāyāṃ ca carataś candrabhāskarau |\\+}
\tl{iḍāyāṃ candramā jñeyaḥ piṃgalāyāṃ raviḥ smṛtaḥ ||\\+}
\tl{candras tāmasa ity uktas sūryo rājasa ucyate |\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_056]
%??JM: rajasic and tamasic not in Eng dictionary, "associated with the \emph{rajas guṇa}" ?
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.57}
\begin{translation}[hp04_057]
Those two bring about the entirety of time, which consists of night and day. Suṣumṇā consumes time. This which has been taught is secret.
\end{translation}

\begin{sources}[hp04_057]
Vasiṣṭhasaṃhitā 2.29cd–30ab
\begin{versinnote}
\tl{tāv eva sakalaṃ dhattaḥ kālaṃ rātrindivātmakam |\\+}
\tl{bhoktrī suṣumṇā kālasya guhyam etad udāhṛtam ||\\!}
\end{versinnote}

Yogayājñavalkya 4.34cd–35ab
\begin{versinnote}
\tl{tāv eva dhattaḥ sakalaṃ kālaṃ rātridivātmakam |\\+}
\tl{bhoktrī suṣumnā kālasya guhyam etad udāhṛtam ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_057]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{tāv eva dhattas sakalaṃ kālaṃ rātriṃ divātmakam |\\+}
\tl{bhoktrī suṣumṇā kālasya guhyam etad udāhṛtam ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} (check in ms. no. 2244)
\begin{versinnote}
\tl{tathā coktaṃ haṭhapradīkāyāṃ–\\+}
\tl{sūryācandramasau dhattaḥ lagnaḥ kāla rātridinātmakaṃ ||\\+}
\tl{bhoktrī suṣumnā kālasya guhyate tad udīritaṃ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_057]
\end{philcomm}

\begin{metre}[hp04_057]
Anuṣṭubh (a: bha-vipulā; c: ma-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.58}
\begin{translation}[hp04_058]
For as the tetrad of verses called the Saubhadra has it:

There are six cakras, sixteen supports, three focal points and three guṇas. Everything else is [just] the prolixity of texts. Trikūṭa is the supreme place.
\end{translation}

%\begin{sources}[hp04_058]
%\end{sources}

\begin{testimonia}[hp04_058]
Cf.~6-chapter \emph{Vivekamārtaṇḍa} 6.3
\begin{versinnote}
\tl{ṣaṭcakraṃ ṣoḍaśādhāraṃ trilakṣaṃ vyomapañcakam |\\+}
\tl{svadehe ye na jānanti kathaṃ sidhyanti yoginaḥ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_058]
We do not know why this tetrad of verses is called Saubhadra.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.59}
\begin{translation}[hp04_059]
Kuṇḍalinī is taught to have a curved shape like a snake. She is Śakti. [The yogi] who has made her move is undoubtedly liberated.
\end{translation}

%\begin{sources}[hp04_059]
%\end{sources}

%\begin{testimonia}[hp04_059]
%\end{testimonia}

%\begin{philcomm}[hp04_059]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.60}
\begin{translation}[hp04_060]
When kūṭa is situated at trikūṭa [then] the mind is wonderful and uninterrupted. By means of kuṇḍalinī, [the yogi] is undoubtedly liberated.
\end{translation}

%\begin{sources}[hp04_060]
%\end{sources}

\begin{testimonia}[hp04_060]
?? Upāsanāsārasaṅgraha !!!
\end{testimonia}

\begin{philcomm}[hp04_060]
We are unsure of the referent of \emph{kūṭa} here. As suggested to us by Sven Sellmer, it may mean the tip of the tongue, which in e.g.~\emph{Khecarīvidyā} 1.65–67 and 3.16–17, is to be placed at \emph{trikūṭa} as part of the practice of \emph{khecarīmudrā}.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.61}
\begin{translation}[hp04_061]
There are seventy-two thousand openings  of the channels in the cage [that is the body]. Suṣumṇā is the Śāmbhavī Śakti while the other [channels] are pointless.
\end{translation}

%\begin{sources}[hp04_061]
%\end{sources}

\begin{testimonia}[hp04_061]
\emph{Haṭhasaṅketacandrikā} f.~108r
\begin{versinnote}
\tl{sūryā[c]andram asau kṛtvā viditvā karapaṃjare [|] \\+}
\tl{suṣumṇā śāṃbhavī śaktiḥ śeṣās tv eva nirarthakāḥ ||\\!}
\end{versinnote}

Yogaśikhopaniṣat 6.17cd–18ab
\begin{versinnote}
\tl{dvisaptatisahasrāṇi nāḍīdvārāṇi pañjare ||\\+}
\tl{suṣumnā śāmbhavī śaktiḥ śeṣās tv anye nirarthakāḥ |\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_061]
The compound \emph{nāḍīdvāra} is not found elsewhere (other than as \emph{nāḍīdvāreṇa}) and its meaning here is unclear. Brahmānanda understands \emph{dvārāṇi} to refer to routes by which breath enters the body (\emph{dvārāṇi vāyupraveśamārgāḥ}) and we have translated \emph{nāḍīdvārāṇi} accordingly.
%Consider °nāḍīnāṃ dehapañjare ??JM: probably a patch?
\end{philcomm}

%%%%%%%%%%
\subsection*{4.62}
\begin{translation}[hp04_062]
The breath, having been carefully accumulated, together with fire awakens kuṇḍalinī and enters Suṣumṇā without obstruction.
\end{translation}

\begin{sources}[hp04_062]
\emph{Dattātreyayogaśāstra} 108
\begin{versinnote}
\tl{vāyuḥ paricito yatnād agninā saha kuṇḍalīm |\\+}
\tl{bodhayitvā suṣumnāyāṃ praviśed avirodhataḥ ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_062]
\emph{Śārṅgadharapaddhati} 4399
\begin{versinnote}
\tl{vāyuḥ paricito yasmād agninā saha kuṇḍalīm  |\\+}
\tl{bodhayitvā suṣumnāyāṃ praviśed anirodhataḥ ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} ff.~197v–180r (attributed to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{vāyuḥ paricito yasmād agninā saha kuṃḍalī |\\+}
\tl{bodhayitvā suṣumnāyāṃ praviśed anirodhata iti\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_062]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.63}
\begin{translation}[hp04_063]
When the breath is flowing in Suṣumṇā, the no-mind state is sure to be attained. Otherwise [i.e.~if the breath is not flowing in Suṣumṇā] the various practices [of yoga] lead to nothing but exertion for yogis.
\end{translation}

%\begin{sources}[hp04_063]
%\end{sources}

\begin{testimonia}[hp04_063]
\emph{Upāsanāsārasaṅgraha} p.~108
\begin{versinnote}
\tl{haṭhapradīpikāyām–\\+}
\tl{suṣumnāvāhinī prāṇe siddhaty eva manonmani |\\+}
\tl{anyathā vividhābhyāso prayāsāyaiva yoginām ||\\!}
\end{versinnote}

\emph{Haṭhasaṅketacandrikā} f.~113v
\begin{versinnote}
\tl{suṣumṇāvāhini prāṇe sidhyaty eva manonmanī  |\\+}
\tl{anye ye vividhābhyāsāḥ prayāsā eva yogināṃ  ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_063]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.64}
\begin{translation}[hp04_064]
The mind is bound by the very same thing that binds the breath and the breath is bound by that which binds the mind.
\end{translation}

%\begin{sources}[hp04_064]
%\end{sources}

\begin{testimonia}[hp04_064]

\emph{Haṭhasaṅketacandrikā} f.~67r
\begin{versinnote}
\tl{tathā coktaṃ haṭhapradīpikāyāṃ–\\+}
\tl{pavano badhyate yena manas tenaiva b[a]dhyate \\+}
\tl{mānaś ca badhyate yena pavanas tena badhyate [|]\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_064]
Brahmānanda understands \emph{yena} here to refer to the yogi. We have taken it to refer to a practice.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.65}
\begin{translation}[hp04_065]
The mind has two impulses: past impression (\emph{vāsanā}) and the breath. When one of those two disappears, both soon disappear.%??JM: tr. vāsanā as plural? Any better ideas?
\end{translation}

\begin{sources}[hp04_065]
\emph{Gorakṣaśataka} 9
\begin{versinnote}
\tl{hetudvayañ ca cittasya vāsanā ca samīraṇaḥ |\\+}
\tl{tayor vinaṣṭa ekasmin drutaṃ dvāv api naśyataḥ |\\+}
\tl{\var{9d drutaṃ dvāv api ] em.; dhṛtaṃ dvāv api T, tasmai dvāv api G1, nasmai dvāv api G2, tad dvāv api vi° U}\\!}
\end{versinnote}

\emph{Mokṣopāya} V.92.48
\begin{versinnote}
\tl{dve bīje rāma cittasya prāṇaspandanavāsane |\\+}
\tl{ekasmiṃś ca tayoẖ kṣīṇe kṣipraṃ dve api naśyataḥ ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_065]
\emph{Yogakuṇḍalinyupaniṣat} 1
\begin{versinnote}
\tl{hetudvayaṃ hi cittasya vāsanā ca samīraṇaḥ |\\+}
\tl{tayor vinaṣṭa ekasmiṃs tad dvāv api vinaśyataḥ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_065]
It is difficult to say whether the \emph{Gorakṣaśataka}'s reading \emph{cittasya} was changed to \emph{manaso} by the author when this verse was borrowed in order to make the terminology consistent with the previous verse, or whether this change occurred at a later time.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.66}
\begin{translation}[hp04_066]
The breath dissolves where the mind dissolves; the mind dissolves right where the breath dissolves.
\end{translation}

%\begin{sources}[hp04_066]
%\end{sources}

\begin{testimonia}[hp04_066]
\emph{Haṭharatnāvalī} 4.29
\begin{versinnote}
\tl{mano yatra vilīyeta pavanas tatra līyate ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_066]
%JM?? do we need this note? I think it's about something other than samādhi.
These verses seem to elevate the possible role of the mind in achieving \emph{samādhi} (as opposed to the standard Haṭha method of stopping the breath to achieve samādhi).

%Brahmānanda understands this verse as referring to specific locations into which the breath and mind dissolve, rather tha"when... then...".
%HI: tatra = tāvat.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.67}
\begin{translation}[hp04_067]
 Like water mixed with milk, mind and breath are always joined, behaving in the same way.  As long as there is mind the breath is active, and as long as there is breath the mind is active.
\end{translation}

\begin{sources}[hp04_067]
\emph{Amanaska} 2.27
\begin{versinnote}
\tl{dugdhāṃbuvat sammilitau sadaiva\\+}
\tl{tulyakriyau mānasamārutau ca |\\+}
\tl{yāvan manas tatra marutpravṛttir \\+}
\tl{yāvan maruc cāpi manaḥpravṛttiḥ |\\+}
\tl{\var{sadaiva ] NI, S, N : tathaiva Cc : sad eva Na}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_067]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{rājayoge—\\+}
\tl{dugdhāmbuvat saṃmilitau sadaiva\\+}
\tl{tulyakriyau mānasamārutau ca |\\+}
\tl{yāvan manas tatra marutpravṛttir \\+}
\tl{yāvan marut tatra manaḥpravṛttiḥ ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 2.5
\begin{versinnote}
\tl{dugdhāmbuvat saṃmilitau sadaiva \\+}
\tl{tulyakriyau mānasamārutau ca | \\+}
\tl{yāvan manas tatra marutpravṛttis \\+}
\tl{tatraikanāśād aparasya nāśaḥ ||\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_067]
Complementing his understanding of the previous verse, Brahmānanda has \emph{yato\dots tatra} in \emph{pāda}s c and d, taking it to mean \emph{yatra\dots tatra} and to be referring to cakras.
\end{philcomm}

\begin{metre}[hp04_067]
Upajāti
\end{metre}

%%%%%%%%%%
\subsection*{4.68}
\begin{translation}[hp04_068]
As a result of one of those two disappearing the other disappears and as a result of one being active the other is active. And when neither has disappeared there is perception through all the sense faculties. When both have disappeared the state of liberation is attained.
\end{translation}

\begin{sources}[hp04_068]
\emph{Amanaska} 2.28
\begin{versinnote}
\tl{tatraikanāśād aparasya nāśa\\+}
\tl{ekapravṛtter aparapravṛttiḥ |\\+}
\tl{adhvastayoś cendriyavargabuddhir \\+}
\tl{vidhvastayor mokṣapadasya siddhiḥ ||\\+}
\tl{\var{vargabuddhir ] NI and S : vargavṛttir Pa Tr Va Nb Ea Eb : vargavṛddhiḥ VbVd: vargavṛddhir N : vargavidhi Pc : sargabuddhir Cc: sargaviddhir Nu}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_068]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{tatraikanāśād aparasya nāśaḥ\\+}
\tl{ekapravṛtter aparapravṛttiḥ |\\+}
\tl{adhvastayoḥ svendriyavargavṛddhir\\+}
\tl{vidhvastayor mokṣapadasya siddhiḥ ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 2.6
\begin{versinnote}
\tl{ekapravṛttāv aparapravṛttir \\+}
\tl{ekasya nāśād aparasya nāśaḥ |\\+}
\tl{adhvastayor indriyavargavṛttir \\+}
\tl{vidhvastayoḥ mokṣapadasya siddhiḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_068]
%Destroy etc. for disappear?
%J5 śuddhir (for buddhir)
%\end{philcomm}

\begin{metre}[hp04_068]
Upajāti
\end{metre}

%%%%%%%%%%
\subsection*{4.69}
\begin{translation}[hp04_069]
[The yogi becomes] a traveller by the way of the wind [and] obtains the whole earth and the eight lordly powers. True, true is this, O beautiful one.
\end{translation}

\begin{sources}[hp04_069]
\emph{Jñānasāra} 3.5–6
\begin{versinnote}
\tl{vāyuvegena deveśi sakalāṃ bhramate mahīm |\\+}
\tl{aṣṭadhāguṇam aiśvaryaṃ satyaṃ satyaṃ na cānyathā ||\\!}
\end{versinnote}
\end{sources}

%\begin{testimonia}[hp04_069]
%\end{testimonia}

%\begin{philcomm}[hp04_069]
%Unidentified source. Jnānasāra removes vocative.
%vāyumārga is an expression for flying (Rāmāyaṇa 7.33.3, etc., etc.)
%labhate needed for aṣṭaguṇaṃ
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.70}
\begin{translation}[hp04_070]
Thus Viśvarūpācārya:\\
When the breath is destroyed and the mind dissolves, all experience is the same. That is called Samādhi.
\end{translation}

\begin{sources}[hp04_070]
\emph{Vivekamārtaṇḍa} 163
\begin{versinnote}
\tl{yadā saṃkṣīyate prāṇo mānasaṃ ca vilīyate |\\+}
\tl{tadā samarasatvaṃ ca samādhiḥ so 'bhidhīyate ||\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_070]
\emph{Yuktabhavadeva} 30
\begin{versinnote}
\tl{yadā saṃkṣīyate prāṇo mānasaṃ ca pralīyate |\\+}
\tl{yadā samarasatvaṃ ca samādhiḥ procyate tadā ||\\+}
\tl{\var{yadā ] tadā}\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_070]
The six-chapter \emph{Vivekamārtaṇḍa} is attributed to Viśvarūpa, which perhaps explains the attribution of this verse to him.
\end{philcomm}

%%%%%%%%%%
\subsection*{4.71}
\begin{translation}[hp04_071]
When the mind is still the breath is still, from which bindu becomes still.  As a result of bindu becoming still, my son, the body becomes still.
\end{translation}

%\begin{sources}[hp04_071]
%\end{sources}

\begin{testimonia}[hp04_071]
\emph{Yogacintāmaṇi}
\begin{versinnote}
\tl{haṭhapradīpikāyām—\\+}
\tl{manaḥsthairye sthiro vāyus tato binduḥ sthiro bhavet |\\+}
\tl{bindusthairyād athāpannaṃ piṇḍasthairyaṃ prajāyata iti ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 43.19
\begin{versinnote}
\tl{cittasthairye mārutasusthiraḥ syāt
tasmād bindususthiro yogino 'ṃge~|\\+}
\tl{bindusthairye syād dayā satvam ojaḥ
piṇḍasthairyaṃ kāyasampad balaṃ ca || 19 ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_071]
%88c āpanna not well attested and N3 has bindusthairyodayāt putra: perhaps this comes from an unknown source text addressed to a son.
%?? putra
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.72}
\begin{translation}[hp04_072]
Only he whose gaze is steady without a visible object, whose breath is steady without effort [and] whose mind is steady without a support is a yogi, is a guru, is to be served.
\end{translation}

\begin{sources}[hp04_072]
\emph{Amanaska} 2.44
\begin{versinnote}
\tl{dṛṣṭiḥ sthirā yasya vinaiva dṛśyād\\+}
\tl{vāyuḥ sthiro yasya vinā prayatnāt |\\+}
\tl{cittaṃ sthiraṃ yasya vināvalambāt\\+}
\tl{sa eva yogī sa guruḥ sa sevyaḥ ||\\+}
\tl{\var{vinaiva dṛśyād  AllNI AllSI Eb: vinā sudṛśyād Ua: vinā nimeṣād Pb AllN : vinaiva lakṣyād Cc: vinaiva dṛṣṭyād Jb: vinakadṛśyad Ma: vinānaṅgame Nb}\\!}
\end{versinnote}
\end{sources}

\begin{testimonia}[hp04_072]
\emph{Yogacintāmaṇi} (attributed to the \emph{Rājayoga})
\begin{versinnote}
\tl{dṛṣṭiḥ sthirā yasya vinaiva dṛśyaṃ\\+}
\tl{vāyuḥ sthiro yasya vinā prayatnaṃ |\\+}
\tl{cittaṃ sthiraṃ yasya vināvalaṃbaṃ\\+}
\tl{sa eva yogī sa guruḥ saṃsevyaḥ ||\\!}
\end{versinnote}

\emph{Kulārṇavatantra} 13.70
\begin{versinnote}
\tl{dṛśyaṃ vinā sthirā dṛṣṭir manaś cālambanaṃ  vinā |\\+}
\tl{vināyāsaṃ sthiro vāyur yasya syāt sa guruḥ priye ||\\!}
\end{versinnote}

\emph{Haṭharatnāvalī} 4.25
\begin{versinnote}
\tl{dṛṣṭiḥ sthirā yasya vinaiva lakṣyāt\\+}
\tl{vāyuḥ sthiro yasya vinā prayatnāt |\\+}
\tl{cittaṃ sthiraṃ yasya vināvalambanāt\\+}
\tl{sa eva yogī sa guruḥ sa sevyaḥ ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 16
\begin{versinnote}
\tl{dṛṣṭiḥ sthirā yasya vinaiva dṛśyād\\+}
\tl{vāyuḥ sthiro yasya vinā prayatnāt |\\+}
\tl{cittaṃ sthiraṃ yasya vinā'valambāt \\+}
\tl{sa eva yogī sa guruḥ sa sevyaḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_072]
%\end{philcomm}

\begin{metre}[hp04_072]
Upajāti
\end{metre}

%%%%%%%%%%
\subsection*{4.73}
\begin{translation}[hp04_073]
[The yogi] whose breath does not move in, out, left, right, up or down is liberated. In this there is no doubt.
\end{translation}

\begin{sources}[hp04_073]
\emph{Gorakṣaśataka} 8
\begin{versinnote}
\tl{praveśe nirgame vāme dakṣiṇe cordhvam apy adhaḥ |\\+}
\tl{na yasya vāyur vrajati sa mukto nātra saṃśayaḥ ||\\!}
\end{versinnote}
\end{sources}

%\begin{testimonia}[hp04_073]
%\end{testimonia}

%\begin{philcomm}[hp04_073]
%\end{philcomm}

\begin{metre}[hp04_073]
Anuṣṭubh (c: bha-vipulā)
\end{metre}

%%%%%%%%%%
\subsection*{4.74}
\begin{translation}[hp04_074]
All the methods of Haṭha and Laya are for mastering Rājayoga. The man who has ascended to Rājayoga cheats death.
\end{translation}

%\begin{sources}[hp04_074]
%\end{sources}

\begin{testimonia}[hp04_074]
\emph{Yogacintāmaṇi} (attrib.~to the \emph{Haṭhapradīpikā})
\begin{versinnote}
\tl{sarve layahaṭhābhyāsā rājayogasya siddhaye |\\+}
\tl{rājayogaṃ samārūḍhaḥ puruṣaḥ kālavañcaka iti ||\\!}
\end{versinnote}

\emph{Haṭhatattvakaumudī} 55.34
\begin{versinnote}
\tl{haṭhapradīpikāyām–\\+}
\tl{sarve haṭhalayopāyā rājayogasya siddhaye |\\+}
\tl{rājayogasamārūḍhaḥ puruṣaḥ kālavañcakaḥ ||\\!}
\end{versinnote}

\emph{Haṃsavilāsa} p.~49
\begin{versinnote}
\tl{sarve haṭhalayopāyā rājayogāya kevalam |\\+}
\tl{rājayogaṃ samārūḍhaḥ puruṣaḥ kālavañcakaḥ ||\\!}
\end{versinnote}
\end{testimonia}

%\begin{philcomm}[hp04_074]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.74*1}
\begin{translation}[hp04_074_1]
Iḍā is the divine Gaṅgā, Piṅgalā is the river Yamunā. Between those two is Suṣumṇā, who is to be recognised as Sarasvatī.
\end{translation}

%\begin{sources}[hp04_074_1]
%\end{sources}

%\begin{testimonia}[hp04_074_1]
%\end{testimonia}

%\begin{philcomm}[hp04_074_1]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.74*2}
\begin{translation}[hp04_074_2]
The place of the Triveṇī confluence is called the king of sacred sites. One should bathe there [and] be freed from all sins.
\end{translation}

%\begin{sources}[hp04_074_2]
%\end{sources}

%\begin{testimonia}[hp04_074_2]
%\end{testimonia}

%\begin{philcomm}[hp04_074_2]
%\end{philcomm}

%%%%%%%%%%
\subsection*{4.75}
\begin{translation}[hp04_075]
O great yogis, experience the nectar of Haṭha, the essence extracted from the ocean of all yoga scriptures after it has thus been churned, if you wish not to grow old and die.
\end{translation}

%\begin{sources}[hp04_075]
%\end{sources}

\begin{testimonia}[hp04_075]
\emph{Haṭhasaṅketacandrikā} f.~145v
\begin{versinnote}
\tl{iti tu sakalayogaśāstrasindhoḥ \\+}
\tl{parimathitād avakṛṣya sārabhūtaṃ [|]\\+}
\tl{anubhavata haṭhāmṛtaṃ yamīndrā\\+}
\tl{yadi bhavatām ajarāmaratvavāṃchā [||]\\!}
\end{versinnote}
\end{testimonia}

\begin{philcomm}[hp04_075]
\end{philcomm}

\begin{metre}[hp04_075]
Puṣpitāgrā
\end{metre}

%%%%%%%%%%
\subsection*{4.75*1}
\begin{translation}[hp04_75_1]
The wise people in the world wash away sin at the sacred site of knowledge (\emph{vidyātīrthe}), the virtuous at the sacred site of truth (\emph{satyatīrthe}), the impure-minded at the sacred site of Gaṅgā (\emph{gaṅgātīrthe}), yogis at the sacred site of knowledge (\emph{jñānatīrthe}), kings at the sacred site of the streams (\emph{dhārātīrthe}), the rich at the sacred site of charity (\emph{dānatīrthe}) [and] women of good family at the sacred site of modesty. (\emph{lajjātīrthe})
\end{translation}

%\begin{sources}[hp04_75_1]
%\end{sources}

\begin{testimonia}[hp04_75_1]
\emph{Yogacintāmaṇi} ??
\end{testimonia}

%\begin{philcomm}[hp04_75_1]
%\end{philcomm}

\begin{metre}[hp04_075_1]
Mandākrāntā 
\end{metre}

\subsection*{Colophon}
\begin{translation}[hp04_col]
Thus ends the fourth chapter in the \emph{Haṭhapradīpikā} composed by the glorious lord amongst yogis Svātmārāma.
\end{translation}

\end{ekdosis}
\end{document}